---
title: "Triple Collocation of VIIRS, AMSR-2, and Buoy Matchups"
output:
  html_notebook:
    toc: True
    theme: united
---


This notebook describes the steps necessary to conduct a triple collocation
study of SST values derived from (a) Suomi_NPP - VIIRS, (b) GCOM-W - AMSR-2 and
(c) in situ observations form moored and drifting buoys.

```{r prep_workspace, include=FALSE}
# Load necessary packages
library(dplyr)
library(ggplot2)
library(mapview)
library(futile.logger)
library(lubridate)
library(tidyr)

ggplot <- function(...) {ggplot2::ggplot(...) + theme_bw()}

# Source own functions
# Define direwctory where functions are for each operating system

if (Sys.info()["sysname"] == 'Windows') {
  fun.dir <- 'D:/matchups/r-projects/Matchup_R_Scripts/Functions/'
} else if (Sys.info()["sysname"] == 'Linux') {
  fun.dir <- '/home/ckk/Projects/Matchup_R_Scripts/Functions/'
}

fun.file <- paste0(fun.dir, 'common_functions.R')

source(file = fun.file,
  local = FALSE, echo = FALSE, verbose = FALSE)
rm(fun.dir, fun.file)
```


# Building a Dataset for Triple Collocation

In a previous step, we read in matchups for Aqua and VIIRS and stored them in
binary (.Rdata) objects. For Aqua, we modified the regular expression to list 
files so that only matchups from 2010 onwards were read. The first step is to
restore those objects. Note that the name of the stored binary objects (.Rdata)
contain the date in which they were created. Be sure to edit the names of the
objects to ensure that the most recent versions are being used.  

```{r load_VIIRS_AQUA_objects, include=FALSE}

# Load .Rdata object for VIIRS matchups resulting from script 01_read_matchups.R
# We assume... TO DO

if (Sys.info()["sysname"] == 'Windows') {
  # Windows directory for VIIRS matchups
  win.dir <- 'D:/matchups/r-projects/R_MUDB/viirs/v641ao/mia/results/objects/'
  # Name of .Rdata file (update as needed)
  win.file <- 'VIIRS_Suomi_NPP_MIA_L2GEN_ALL_Class_6.4.1_ao_2016_11_23_with_ancillary.Rdata'
  viirs.file <- paste0(win.dir, win.file)
  if (!file.exists(viirs.file)) {
    stop('Input file does not exist')
  } else {
    load(viirs.file, verbose = TRUE)
  }
  rm(win.dir, win.file, viirs.file)
} else if (Sys.info()["sysname"] == 'Linux') {
  # Linux directory for VIIRS matchups
  linux.dir <- '~/Projects/Matchup_R_Scripts/Results/objects/'
  # Name of .Rdata file (update as needed)
  linux.file <- 'VIIRS_Suomi_NPP_MIA_L2GEN_ALL_Class_6.4.1_ao_2016_11_22_with_ancillary.Rdata'
  viirs.file <- paste0(linux.dir, linux.file)
  if (!file.exists(viirs.file)) {
    stop('Input file does not exist')
  } else {
    load(viirs.file, verbose = TRUE)
  }
  rm(linux.dir, linux.file, viirs.file)
}

# Create 'VIIRS' object so it has a shorter name
VIIRS <- VIIRS_Suomi_NPP_MIA_L2GEN_ALL_Class_6.4.1_ao_2016_11_22
rm(VIIRS_Suomi_NPP_MIA_L2GEN_ALL_Class_6.4.1_ao_2016_11_22)

# Load .Rdata object for AQUA matchups resulting from script 01_read_matchups.R

if (Sys.info()["sysname"] == 'Windows') {
  # Windows directory for AQUA matchups
  win.dir <- 'D:/matchups/r-projects/R_MUDB/modis/aqua/gsfc/l6cv6/results/objects/'
  # Name of .Rdata file (update as needed)
  win.file <- 'MODIS_Aqua_GSFC_ALL_Class_6.4.1_ao_2016_11_23_with_ancillary.Rdata'
  viirs.file <- paste0(win.dir, win.file)
  if (!file.exists(viirs.file)) {
    stop('Input file does not exist')
  } else {
    load(viirs.file, verbose = TRUE)
  }
  rm(win.dir, win.file, viirs.file)
} else if (Sys.info()["sysname"] == 'Linux') {
  # Linux directory for VIIRS matchups
  linux.dir <- '~/Projects/Matchup_R_Scripts/Results/objects/'
  # Name of .Rdata file (update as needed)
  #linux.file <- 'MODIS_Aqua_GSFC_ALL_Class_6.4.1_ao_2016_11_26_with_ancillary.Rdata'
  linux.file <- 'MODIS_Aqua_GSFC_ALL_Class_6.4.1_ao_2017_03_11_with_ancillary.Rdata'
  aqua.file <- paste0(linux.dir, linux.file)
  if (!file.exists(aqua.file)) {
    stop('Input file does not exist')
  } else {
    load(aqua.file, verbose = TRUE)
  }
  rm(linux.dir, linux.file, aqua.file)
}

# Create 'AQUA' object so it has a shorter name
AQUA <- MODIS_Aqua_GSFC_ALL_Class_6.4.1_ao_2017_03_11
rm(`MODIS_Aqua_GSFC_ALL_Class_6.4.1_ao_2017_03_11`)
```

