---
title: "Triple Collocation of VIIRS, Aqua, and Buoy Matchups"
output:
  html_notebook: default
  html_document: default
---

This notebook describes the steps necessary to conduct a triple collocation
study of SST values derived from (a) AQUA-MODIS, (b) Suomi_NPP-VIIRS and
(c) in situ observations form moored and drifting buoys.

```{r prep_workspace, include=FALSE}
# Load necessary packages
library(dplyr)
library(ggplot2)
library(mapview)
library(futile.logger)
library(lubridate)
library(tidyr)

ggplot <- function(...) {ggplot2::ggplot(...) + theme_bw()}

# Source own functions
# Define direwctory where functions are for each operating system

if (Sys.info()["sysname"] == 'Windows') {
  fun.dir <- 'D:/matchups/r-projects/Matchup_R_Scripts/Functions/'
} else if (Sys.info()["sysname"] == 'Linux') {
  fun.dir <- '/home/ckk/Projects/Matchup_R_Scripts/Functions/'
}

fun.file <- paste0(fun.dir, 'common_functions.R')

source(file = fun.file,
  local = FALSE, echo = FALSE, verbose = FALSE)
rm(fun.dir, fun.file)

```


# Building a Dataset for Triple Collocation

In a previous step, we read in matchups for AQUA and VIIRS and stored them in
binary (.Rdata) objects. For Aqua, we modified the regular expression to list files so that only matchups from 2010 onwards were read.
The first step is to restore those objects.
Note that the name of the stored binary objects (.Rdata) contain the date in which they were created. Be sure to edit the names of the objects to ensure that the most recent versions are being used.  

```{r load_VIIRS_AQUA_objects, include=FALSE}

# Load .Rdata object for VIIRS matchups resulting from script 01_read_matchups.R

if (Sys.info()["sysname"] == 'Windows') {
  # Windows directory for VIIRS matchups
  win.dir <- 'D:/matchups/r-projects/R_MUDB/viirs/v641ao/mia/results/objects/'
  # Name of .Rdata file (update as needed)
  win.file <- 'VIIRS_Suomi_NPP_MIA_L2GEN_ALL_Class_6.4.1_ao_2016_11_23_with_ancillary.Rdata'
  viirs.file <- paste0(win.dir, win.file)
  if (!file.exists(viirs.file)) {
    stop('Input file does not exist')
  } else {
    load(viirs.file, verbose = TRUE)
  }
  rm(win.dir, win.file, viirs.file)
} else if (Sys.info()["sysname"] == 'Linux') {
  # Linux directory for VIIRS matchups
  linux.dir <- '~/Projects/Matchup_R_Scripts/Results/objects/'
  # Name of .Rdata file (update as needed)
  linux.file <- 'VIIRS_Suomi_NPP_MIA_L2GEN_ALL_Class_6.4.1_ao_2016_11_22_with_ancillary.Rdata'
  viirs.file <- paste0(linux.dir, linux.file)
  if (!file.exists(viirs.file)) {
    stop('Input file does not exist')
  } else {
    load(viirs.file, verbose = TRUE)
  }
  rm(linux.dir, linux.file, viirs.file)
}

# Create 'VIIRS' object so it has a shorter name
VIIRS <- VIIRS_Suomi_NPP_MIA_L2GEN_ALL_Class_6.4.1_ao_2016_11_23
rm(VIIRS_Suomi_NPP_MIA_L2GEN_ALL_Class_6.4.1_ao_2016_11_23)

# Load .Rdata object for AQUA matchups resulting from script 01_read_matchups.R

if (Sys.info()["sysname"] == 'Windows') {
  # Windows directory for AQUA matchups
  win.dir <- 'D:/matchups/r-projects/R_MUDB/modis/aqua/gsfc/l6cv6/results/objects/'
  # Name of .Rdata file (update as needed)
  win.file <- 'MODIS_Aqua_GSFC_ALL_Class_6.4.1_ao_2016_11_23_with_ancillary.Rdata'
  viirs.file <- paste0(win.dir, win.file)
  if (!file.exists(viirs.file)) {
    stop('Input file does not exist')
  } else {
    load(viirs.file, verbose = TRUE)
  }
  rm(win.dir, win.file, viirs.file)
} else if (Sys.info()["sysname"] == 'Linux') {
  # Linux directory for VIIRS matchups
  linux.dir <- '~/Projects/Matchup_R_Scripts/Results/objects/'
  # Name of .Rdata file (update as needed)
  linux.file <- 'MODIS_Aqua_GSFC_ALL_Class_6.4.1_ao_2016_11_26_with_ancillary.Rdata'
  aqua.file <- paste0(linux.dir, linux.file)
  if (!file.exists(aqua.file)) {
    stop('Input file does not exist')
  } else {
    load(aqua.file, verbose = TRUE)
  }
  rm(linux.dir, linux.file, aqua.file)
}

# Create 'AQUA' object so it has a shorter name
AQUA <- MODIS_Aqua_GSFC_ALL_Class_6.4.1_ao_2016_11_23
rm(MODIS_Aqua_GSFC_ALL_Class_6.4.1_ao_2016_11_23)

```

The restored VIIRS matchups span the period between `r min(VIIRS$buoy.timedate)` and `r max(VIIRS$buoy.timedate)`.The original VIIRS dataset had `r nrow(VIIRS)` matchups.
We also read Aqua matchups for the period between `r min(AQUA$buoy.timedate)`
and `r max(AQUA$buoy.timedate)`. The number of Aqua matchups was `r nrow(AQUA)`.

These restored objects have many more variables than we need.
The AQUA matchups have `r ncol(AQUA)` variables, and the VIIRS matchups have
`r ncol(VIIRS)` columns. Therefore, a second step eliminates many variables
or columns that are not needed for the collocation study.

```{r filter_variables, include = FALSE}

# Define variables that will be retained for collocation study for each sensor
# and matchup format

modis.colloc.vars <- keep_matchup_variables_collocation(matchup.format = 'GSFC',
  sensor = 'MODIS', ancillary.data = TRUE)

viirs.colloc.vars <- keep_matchup_variables_collocation(matchup.format = 'MIA_L2GEN',
  sensor = 'VIIRS', ancillary.data = TRUE)

AQUA.trimmed <- dplyr::select(AQUA, one_of(modis.colloc.vars))
VIIRS.trimmed <- dplyr::select(VIIRS, one_of(viirs.colloc.vars))

rm(modis.colloc.vars, viirs.colloc.vars)

```

After the reduction of unnecessary variables, the trimmed AQUA matchups
have `r ncol(AQUA.trimmed)` variables, and the trimmed VIIRS matchups have
`r ncol(VIIRS.trimmed)` columns. 

The third preparatory step is to convert the VIIRS BTs and SSTs from degrees Kelvin to Celsius, whereas Aqua BTs, and SSTs already are expressed in Celsius. For consistency we  convert VIIRS values to degrees Celsius from degrees Kelvin.

```{r convert_VIIRS_K_to_C, echo = FALSE}

# Convert VIIRS BTs and SSTs from degrees Kelvin to Celsius

# Identify columns of 'VIIRS.trimmed' that need to be converted from K to C
pattern <- paste("^cen", "^med", "^max", "^min", sep = "|")
tt1 <- which((grepl(pattern, names(VIIRS.trimmed))))

# Convert selected columns from K to C
VIIRS.trimmed[tt1] <- apply(VIIRS.trimmed[tt1], 2, FUN = degK.to.degC)

rm(pattern, tt1)

```

To confirm that the conversion from Kelvin to Celsius was successful, we output the summary of the 'cen.sst' column before and after the conversion.

```{r verify_K_to_C, echo = FALSE}
cat('BEFORE K to C conversion\n')
summary(VIIRS$cen.sst)
cat('AFTER K to C conversion\n')
summary(VIIRS.trimmed$cen.sst)

```

Because VIIRS and Aqua SSTs are calculated as 'skin' values, we neeed to add back 0.17 degC
to each of these SSTs to make them compatible with the buoys' 'bulk' SST.

```{r skin_to_bulk}

VIIRS.trimmed$cen.sst <- VIIRS.trimmed$cen.sst + 0.17
AQUA.trimmed$cen.sst <- AQUA.trimmed$cen.sst + 0.17

```

The final preparatory step before joining the VIIRS and AQUA matchups is to add
a suffix to the variables of each data set indicating the source of the variable. For example, variable 'cen.sst' will be re named to 'cen.sst.vrs' if it corresponds to VIIRS and 'cen.sst.aqu' if it comes from the AQUA-MODIS matchups. We do not change the names of columns 'buoy.id' and 'buoy.date', as those are the common
variables in both datasets that will be used to join them in the next section.

```{r rename_columns, echo=FALSE}

# Change column names of VIIRS and AQUA objects so that when we join
# the source of each variable (VIIRS or AQUA) is clear.

colnames(AQUA.trimmed) <- paste0(colnames(AQUA.trimmed), ".aqu")
colnames(VIIRS.trimmed) <- paste0(colnames(VIIRS.trimmed), ".vrs")

# We do not change the names of columns 'buoy.id' and 'buoy.date'
# because they are used as common keys for joining VIIRS and AQUA data.

tt1 <- colnames(AQUA.trimmed)
tt2 <- which(tt1 %in% "buoy.date.aqu")
colnames(AQUA.trimmed)[tt2] <- "buoy.date"
tt3 <- which(tt1 %in% "buoy.id.aqu")
colnames(AQUA.trimmed)[tt3] <- "buoy.id"

tt1 <- colnames(VIIRS.trimmed)
tt2 <- which(tt1 %in% "buoy.date.vrs")
colnames(VIIRS.trimmed)[tt2] <- "buoy.date"
tt3 <- which(tt1 %in% "buoy.id.vrs")
colnames(VIIRS.trimmed)[tt3] <- "buoy.id"

rm(tt1,tt2,tt3)

# Output a few variables to confirm the renaming
colnames(VIIRS.trimmed[1:10])
colnames(AQUA.trimmed[1:10])

```

## Joining VIIRS and Aqua Matchups

A collocation study requires data from at least three sources that are relatively coincident in space and time.
We require that each collocated record include SST measurements from all three
techniques collected within at most 60 minutes and one kilometer from each other.
We have two SST measurements derived from satellite-based infrared radiometers (Suomi Npp - VIIRS and Aqua - MODIS). Since the buoy data are common elements to both VIIRS and Aqua matchups, we use them as a common key to join Aqua and VIIRS matchups. 

```{r join_VIIRS_and_AQUA, echo=TRUE}

# Convert timedates from POSIXct objects to characters for the dplyr joining
# dplyr does not accept POSIXct variables for joining

VIIRS.trimmed$buoy.timedate.vrs <- as.character(VIIRS.trimmed$buoy.timedate.vrs)
VIIRS.trimmed$sat.timedate.vrs <- as.character(VIIRS.trimmed$sat.timedate.vrs)
AQUA.trimmed$buoy.timedate.aqu <- as.character(AQUA.trimmed$buoy.timedate.aqu)
AQUA.trimmed$sat.timedate.aqu <- as.character(AQUA.trimmed$sat.timedate.aqu)

# Join VIIRS and AQUA into one object (orig_j)

orig_j <- dplyr::inner_join(VIIRS.trimmed, AQUA.trimmed,
  by = c("buoy.date", "buoy.id"))

# This line excludes non-exact matches and we lose too many potentially useful matchups
#orig_j <- dplyr::inner_join(AQUA, VIIRS, by = c("buoy.pftime", "buoy.lon", "buoy.lat", "buoy.id"))

```

```{r dates_back_to_POSIX, echo=TRUE}

# After join, we convert back the timedate objects
# from character into POSIXct objects so they can be used in time calculations

tt1 <- orig_j %>% dplyr::select(buoy.timedate.aqu)
tt2 <- as.vector(tt1[[1]])
tt3 <- lubridate::ymd_hms(tt2)
orig_j$buoy.timedate.aqu <- tt3

tt1 <- orig_j %>% dplyr::select(sat.timedate.aqu)
tt2 <- as.vector(tt1[[1]])
tt3 <- lubridate::ymd_hms(tt2)
orig_j$sat.timedate.aqu <- tt3

tt1 <- orig_j %>% dplyr::select(buoy.timedate.vrs)
tt2 <- as.vector(tt1[[1]])
tt3 <- lubridate::ymd_hms(tt2)
orig_j$buoy.timedate.vrs <- tt3

tt1 <- orig_j %>% dplyr::select(sat.timedate.vrs)
tt2 <- as.vector(tt1[[1]])
tt3 <- lubridate::ymd_hms(tt2)
orig_j$sat.timedate.vrs <- tt3

rm(tt1,tt2,tt3)
```

## Filtering VIIRS and Aqua Matchups

At this point, we have `r nrow(orig_j)` joined matchups. However, these matchups do not have enough spatial or temporal coincidence to meet our requirements stated above or for a triple collocation error analysis. We did not join by the buoy's timedate or latitude/longitude because the 'dplyr::inner_join' function only matches if key values are *exactly* the same in both datasets, thereby reducing the number of potentially useful matchups. We instead apply a series of filtering steps to ensure the conditions stated above for a triple collocation error analysis.
We also filter matchups that have good quality so that we minimize chances of
errors in satellite SST retrievals due to cloud contamination.

```{r filtering_orig_j, echo=FALSE}

# Filter collocated data by solz, platform, time difference

orig_filtered <- dplyr::tbl_df(orig_j) %>%
  # Only nightime retrievals (solar zenith angle >= 90 degrees)
  dplyr::filter(solz.aqu >= 90 & solz.vrs >= 90) %>%
  # Allow only buoy SST measurements (not ships or radiometers) 
  dplyr::filter(insitu.platform.aqu %in% c("DriftingBuoy","MooredBuoy")) %>%
  # Eliminate SST qualities for AQUA and VIIRS less than 0 (best quality) 
  dplyr::filter(qsst.aqu == 0 & qsst.vrs == 0) %>%
  # VIIRS and AQUA sat times must be less or equal than 1hr (3600 secs) from each other
  dplyr::filter(abs(sat.timedate.aqu - sat.timedate.vrs) <= 3600) %>%
  # Buoy retrieval times for VIIRS and AQUA must be less than 1 min apart
  # The buoy times should coincide but may have roundoff differences
  dplyr::filter(abs(buoy.timedate.aqu - buoy.timedate.vrs) <= 60)

# Allow only retrievals for which distance between the buoy positions
# listed in VIIRS and AQUA are less than 100 meters

uu1 <- cbind(orig_filtered$buoy.lon.vrs, orig_filtered$buoy.lat.vrs)
uu2 <- cbind(orig_filtered$buoy.lon.aqu, orig_filtered$buoy.lat.aqu)
uu3 <- geosphere::distVincentyEllipsoid(uu1, uu2)
cat('Distance between buoy positions in VIIRS and AQUA\n')
summary(uu3)
uu4 <- which(uu3 < 100) # Identify distances < 100 m
orig_filtered <- orig_filtered[uu4, ] # Filter out distances >= 100 m
rm(uu1,uu2,uu3,uu4)

# Explore distances between satellite and buoy positions for VIIRS
uu1 <- cbind(orig_filtered$buoy.lon.vrs, orig_filtered$buoy.lat.vrs)
uu2 <- cbind(orig_filtered$sat.lon.vrs, orig_filtered$sat.lat.vrs)
uu3 <- geosphere::distVincentyEllipsoid(uu1, uu2)
cat('Distance between sat and buoy positions for VIIRS\n')
summary(uu3)
rm(uu1,uu2,uu3)

# Explore distances between satellite and buoy positions for AQUA
uu1 <- cbind(orig_filtered$buoy.lon.aqu, orig_filtered$buoy.lat.aqu)
uu2 <- cbind(orig_filtered$sat.lon.aqu, orig_filtered$sat.lat.aqu)
uu3 <- geosphere::distVincentyEllipsoid(uu1, uu2)
cat('Distance between sat and buoy positions for AQUA\n')
summary(uu3)
rm(uu1,uu2,uu3)

# Explore distances between VIIRS and AQUA pixel positions
uu1 <- cbind(orig_filtered$sat.lon.vrs, orig_filtered$sat.lat.vrs)
uu2 <- cbind(orig_filtered$sat.lon.aqu, orig_filtered$sat.lat.aqu)
uu3 <- geosphere::distVincentyEllipsoid(uu1, uu2)
cat('Distance between satellite pixel positions for AQUA and VIIRS\n')
summary(uu3)
rm(uu1,uu2,uu3)

```

After applying the various filters we are left with `r nrow(orig_j)`
matchups.

## Spatial distribution of collocated matchups

```{r points_on_map}
library(maps)
library(mapdata)

CairoSVG(file = 'matchups_map.ps',
         width = 9, height = 6, onefile = TRUE,
         bg = "transparent", pointsize = 18)

maps::map("world",
  col = "gray70",
  fill = TRUE)

points(orig_filtered$buoy.lon.aqu, orig_filtered$buoy.lat.aqu,
  pch = '.', col = 'tomato')
box()
dev.off()

```

```{r points_on_map_ggmap}
# Plot location of collocated data 

library(ggmap)
mp <- NULL
mapWorld <- borders("world", colour="gray50", fill="gray70") # create a layer of borders
mp <- ggplot() +   mapWorld
mp <- mp + ggplot2::labs(x = NULL, y = NULL)

# Now layer the matchup locations on top
mp <- mp + ggplot2::geom_point(data = orig_filtered,
  aes(x = buoy.lon.aqu, y = buoy.lat.aqu),
  color = "tomato", size = 0.7, alpha = 0.2)
ggplot2::ggsave(filename = 'point_distribution.ps', device = 'ps',
       width = 8, height = 6, units = 'in')


+
ggplot2::scale_x_continuous(breaks = seq(from = -180, to = 180, by = 60)) +
ggplot2::scale_y_continuous(breaks = seq(from = -90, to = 90, by = 30)) +


plot(mp)

```

```{r map_with_hexbinning}

# Create base map
require(ggmap)

mp <- NULL
mapWorld <- borders("world", colour = "gray50", fill = "gray70") # create a layer of borders
mp <- ggplot() +  mapWorld +
  ggplot2::scale_x_continuous(breaks = seq(from = -180, to = 180, by = 60)) +
  ggplot2::scale_y_continuous(breaks = seq(from = -90, to = 90, by = 30)) +
  coord_fixed(ratio = 1)


# Now layer the buoys on top with hexagonal binning
mp <- mp + 
  ggplot2::stat_binhex(data = orig_filtered,
    aes(x = buoy.lon.aqu, # Use hexagonal binning command and give x and y input
    y = buoy.lat.aqu,
    fill = cut(..value.., c(0, 100, 500, 1000, 1500, Inf))), # Divides matchups per bin into discrete chunks and colors likewise
    binwidth = c(10, 10)) +
  mapWorld + labs(x = NULL, y = NULL) +
  #scale_fill_hue('value') + # Standard colors with discrete chunking
  scale_fill_brewer(palette = 'YlOrRd') + # Change colors to Yellow, Orange, and Red - many diff 
  guides(fill = guide_legend(title = "N of matchups"))

ggplot2::ggsave(filename = 'point_distribution.ps', device = 'ps',
       width = 8, height = 6, units = 'in')

```

## Temporal distribution of collocated matchups

```{r heatmap_of_matchups}

yy <- lubridate::year(orig_filtered$buoy.timedate.vrs)     # Year
mm <- lubridate::month(orig_filtered$buoy.timedate.vrs)    # Month

tt1 <- cbind(yy, mm)

tt2 <- dplyr::tbl_df(tt1) %>%
  dplyr::group_by(yy, mm) %>%
  dplyr::summarise(n = n()) %>%
  dplyr::arrange(yy, mm)

yy.vals <- sort(unique(yy))
mm.vals <- sort(unique(mm))
tt3 <- expand.grid(mm = mm.vals, yy = yy.vals)

tt4 <- dplyr::left_join(tt3, tt2) %>%
  dplyr::select(yy, mm, n)

brks <- pretty(tt4$n, n = 5)
#brks <- floor(quantile(tt4$n, probs = seq(0, 1, 0.25), na.rm = TRUE))

tt4$n <- cut(tt4$n, breaks = brks)
tt4$mm <- ordered(tt4$mm, labels = month.abb)
tt4$yy <- ordered(tt4$yy)

hm <- ggplot2::ggplot(data = tt4, aes(mm, yy)) +
  ggplot2::geom_tile(aes(fill = n), colour = "white") +
  ggplot2::scale_fill_brewer(type = "seq", palette = 'YlOrRd', direction = 1) +
  ggplot2::theme_bw() +
  ggplot2::theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  ggplot2::ggsave("Temporal_heatmap.ps", device = 'ps',
       width = 8, height = 6, units = 'in')



```

## Association between all SST sources

```{r viirs_aqua_sst, echo=FALSE}

ccc <- RColorBrewer::brewer.pal(n = 8, name = 'YlOrRd')
ddd <- orig_filtered$cen.sst.aqu - orig_filtered$cen.sst.vrs

ggplot(data = orig_filtered, mapping = aes(x = cen.sst.aqu, y = ddd)) +
  stat_bin2d(binwidth = c(0.2, 0.1)) +
  scale_fill_gradientn(colours = ccc) +
  geom_abline(intercept = 0, slope = 0, color = 'steelblue') +
  lims(x = c(-2, 32), y = c(-3,3)) +
  ggplot2::labs(title = "VIIRS and AQUA SSTs", x = 'AQUA SST', y = 'VIIRS SST minus AQUA SST') +
  ggplot2::ggsave(filename = 'AQUA_and_VIIRS.ps', device = 'ps',
       width = 8, height = 6, units = 'in') +
  theme(text = element_text(size = 16))

fit_aqua_viirs_sst <- lm(cen.sst.aqu ~ cen.sst.vrs, data = orig_filtered)

# Compute robust standard deviation

RSD_aqua_viirs <- IQR(fit_aqua_viirs_sst$residuals) / 1.348

```

```{r viirs_buoy_sst, echo=FALSE}
ccc <- RColorBrewer::brewer.pal(n = 8, name = 'YlOrRd')
ddd <- orig_filtered$cen.sst.vrs - orig_filtered$buoy.sst.vrs

ggplot(data = orig_filtered, mapping = aes(x = cen.sst.vrs, y = ddd)) +
  stat_bin2d(binwidth = c(0.2, 0.1)) +
  scale_fill_gradientn(colours = ccc) +
  geom_abline(intercept = 0, slope = 0, color = 'steelblue') +
  lims(x = c(-2, 32), y = c(-3,3)) +
  ggplot2::labs(title = "VIIRS and Buoy SSTs", x = 'VIIRS SST', y = 'VIIRS SST minus Buoy SST') +
  ggplot2::ggsave(filename = 'VIIRS_and_Buoy.ps', device = 'ps',
       width = 8, height = 6, units = 'in') +
  theme(text = element_text(size = 16))

fit_viirs_buoy_sst <- lm(cen.sst.vrs ~ buoy.sst.vrs, data = orig_filtered)

# Compute robust standard deviation

RSD_viirs_buoy <- IQR(fit_viirs_buoy_sst$residuals) / 1.348
```

```{r aqua_buoy_sst, echo=FALSE}
ccc <- RColorBrewer::brewer.pal(n = 8, name = 'YlOrRd')
ddd <- orig_filtered$cen.sst.aqu - orig_filtered$buoy.sst.vrs

ggplot(data = orig_filtered, mapping = aes(x = cen.sst.aqu, y = ddd)) +
  stat_bin2d(binwidth = c(0.2, 0.1)) +
  scale_fill_gradientn(colours = ccc) +
  geom_abline(intercept = 0, slope = 0, color = 'steelblue') +
  lims(x = c(-2, 32), y = c(-3,3)) +
  ggplot2::labs(title = "AQUA and Buoy SSTs", x = 'AQUA SST', y = 'AQUA SST minus Buoy SST') +
  ggplot2::ggsave(filename = 'AQUA_and_Buoy.ps', device = 'ps',
       width = 8, height = 6, units = 'in') +
  theme(text = element_text(size = 16))

fit_aqua_buoy_sst <- lm(cen.sst.aqu ~ buoy.sst.vrs, data = orig_filtered)

# Compute robust standard deviation

RSD_aqua_buoy <- IQR(fit_aqua_buoy_sst$residuals) / 1.348
```

# Error estimation using triple collocation

## Global experiment: all matchups in collocated DB (CDB)

```{r collocation_functions, include=FALSE}

compute_bias <- function(source1, source2, source3) {
  if (!is.null(nrow(source1)) |
    !is.null(nrow(source2)) |
    !is.null(nrow(source3)) |
    length(source1) <= 1 |
    length(source2) <= 1 |
    length(source3) <= 1) {
    stop("Only vectors can be passed as arguments to the calculate_bias function.")
  }

  bias_source1source3 <- mean(source1) - mean(source3) #Bias between source1 and source2
  
  bias_source2source3 <- mean(source2) - mean(source3) #Bias between source2 and source3
  
  # Return a dataframe of biases that can re-scale each observation type
  
  biases <- c(bias_source1source3, bias_source2source3)
  names(biases) <- c("bias_src1src3", "bias_src2src3")
  
  return(biases)
}   # End of compute_bias


compute_variance <- function(source1, source2, source3) {
  if (!is.null(nrow(source1)) |
    !is.null(nrow(source2)) |
    !is.null(nrow(source3)) |
    length(source1) <= 1 |
    length(source2) <= 1 |
    length(source3) <= 1) {
    stop("Only vectors can be passed as arguments to the calculate_variance function")
  }
  
  # Call bias function to re-scale measurements
  biases <- compute_bias(source1, source2, source3)
  
  #Scale sources to one source to compute variance
  source1_scaled <- source1 - biases['bias_src1src3']
  source2_scaled <- source2 - biases['bias_src2src3']
  
  # Variance of errors in observation types
  variance_source1 <- mean((source1_scaled - source2_scaled) * (source1_scaled - source3))              # Variance of source1
  variance_source2 <- mean((source1_scaled - source2_scaled) * (source2_scaled - source3))              # Variance of source2
  variance_source3 <- mean((source1_scaled - source3) * (source2_scaled - source3))                     # Variance of source3
  
  # Take all the individual variances and return them as a dataframe
  variances <- data.frame(variance_source1, variance_source2, variance_source3)
  colnames(variances) <- c("variance_source1", "variance_source2", "variance_source3")
  return(variances)
} # end of compute_variance

```

```{r run_collocation_global}

# Make sure the last argument to calculate_variance corresponds to the buoy SSTs

variances_global <- compute_variance(orig_filtered$cen.sst.aqu,
  orig_filtered$cen.sst.vrs,
  orig_filtered$buoy.sst.vrs)

biases_global <- compute_bias(orig_filtered$cen.sst.aqu,
  orig_filtered$cen.sst.vrs,
  orig_filtered$buoy.sst.vrs)

sd_global <- round(sqrt(abs(variances_global)), 3)

```

```{r aqua_viirs_errors}

tt1 <- orig_filtered$cen.sst.aqu - orig_filtered$buoy.sst.vrs
tt2 <- orig_filtered$cen.sst.vrs - orig_filtered$buoy.sst.vrs
ccc <- RColorBrewer::brewer.pal(n = 9, name = 'YlOrRd')[3:9]

errors <- ggplot(mapping = aes(x = tt1, y = tt2)) +
  stat_bin2d(binwidth = c(0.1, 0.1)) +
  scale_fill_gradientn(colours = ccc) +
  geom_abline(intercept = 0, slope = 1, color = 'steelblue') +
  lims(x = c(-3, 3), y = c(-3,3)) +
  ggplot2::labs(title = "AQUA and VIIRS SST Errors", x = 'AQUA SST minus Buoy SST', y = 'VIIRS SST minus Buoy SST') +
  ggplot2::ggsave(filename = 'AQUA_and_VIIRS_errors.ps', device = 'ps',
       width = 8, height = 8, units = 'in') +
  theme(text = element_text(size = 16))

fit_aqua_viirs_errors <- lm(tt1 ~ tt2)

# Compute robust standard deviation

RSD_aqua_viirs_errors <- IQR(fit_aqua_viirs_errors$residuals) / 1.348

```

```{r  colloc_corr_assumptions_sim, echo=FALSE}

# Create an empirical curve for the over/under-estimation of variances with varying correlations

# Set seed so findings are reproducible
set.seed(2)

nobs <- 65000 # Number of observations
ntrials <- 100.00 # Number of trials
corr_values <- seq(from = 0, to = .99, by = .01) # Increments of correlation to test

# Truth variable
truthx <- runif(n = nobs, min = 0, max = 30)

# Biases
alpha1 <- 0
alpha2 <- 0
alpha3 <- 0

# SDs for each observation source
sd_aq <- .4
sd_vi <- .4 # sd_aq should roughly equal sd_vi
sd_bu <- .25

# Dataframe to store predicted variances per correlation
preds_case <- as.data.frame(matrix(999, ncol = 3, nrow = length(corr_values)))
colnames(preds_case) <- c("meas1", "meas2", "meas3")
rownames(preds_case) <- as.character(corr_values)

for (i in 1:length(corr_values)) {
  # Create correlation matrix
  R <- matrix(cbind(1.0,corr_values[i],0.0,  corr_values[i],1.0,0.0,  0.0,0.0,1.0),nrow=3)
  U <- t(chol(R))
  
  # Create dataframe to store predicted variances
  preds_trials <- as.data.frame(matrix(999, ncol = 3, nrow = ntrials))
  colnames(preds_trials) <- c("Meas1", "Meas2", "Meas3")
  rownames(preds_trials) <- as.character(seq(from = 1, to = ntrials, by = 1))
  
  for (trials in seq(from = 1, to = ntrials, by = 1)) {
    if (trials %in% c(1, 50, 100)) {
      cat('Trial', trials, 'of correlation value', corr_values[i], '\n')
    }
    # Create uncorrelated errors
    noi1 <- rnorm(nobs, 0, sd_aq)
    noi2 <- rnorm(nobs, 0, sd_vi)
    noi3 <- rnorm(nobs, 0, sd_bu)
    
    # Correlate the uncorrelated errors
    random.normal <- t(cbind(noi1, noi2, noi3))
    X <- U %*% random.normal
    newX <- t(X)
    raw <- as.data.frame(newX)
    
    # Correlated errors
    epsi1 <- raw[, 1]
    epsi2 <- raw[, 2]
    epsi3 <- raw[, 3]
    
    # Create 3 datasets with biases and correlated errors
    meas1 <- truthx + alpha1 + epsi1
    meas2 <- truthx + alpha2 + epsi2
    meas3 <- truthx + alpha3 + epsi3
    
    # Compute predicted variances
    variances <- round(sqrt(abs(compute_variance(meas1, meas2, meas3))), 3)
    
    # Store predicted variances in a dataframe
    preds_trials[trials, ] <- variances
  }
  # Take the median of each dataset's predicted variance and store in a dataframe
  cat('Taking the median of all trials for correlation value', corr_values[i], '\n')
  #if (corrs == .29) {
  #  preds_case[30, 1] <- median(preds_trials[, 1])
  #  preds_case[30, 2] <- median(preds_trials[, 2])
  #  preds_case[30, 3] <- median(preds_trials[, 3])
  #} else if(corrs == .58) {
  #  preds_case[59, 1] <- median(preds_trials[, 1])
  #  preds_case[59, 2] <- median(preds_trials[, 2])
  #  preds_case[59, 3] <- median(preds_trials[, 3])
  #} else {
  preds_case[i, 1] <- median(preds_trials[, 1])
  preds_case[i, 2] <- median(preds_trials[, 2])
  preds_case[i, 3] <- median(preds_trials[, 3])
  #}
}

# Prep dataframe for plotting
prop1 <- preds_case[, 1] / preds_case[1, 1]
prop2 <- preds_case[, 2] / preds_case[1, 2]
prop3 <- preds_case[, 3] / preds_case[1, 3]

# Linear models fitting the proportions as a function of correlation
fit_prop1 <- lm(prop1 ~ corr_values)
fit_prop2 <- lm(prop2 ~ corr_values)
fit_prop3 <- lm(prop3 ~ corr_values)

int1 <- fit_prop1$coefficients[1]
slo1 <- fit_prop1$coefficients[2]

int2 <- fit_prop2$coefficients[1]
slo2 <- fit_prop2$coefficients[2]

# Plot proportions
corr_sim <- ggplot2::ggplot() +
  geom_line(aes(corr_values, prop1), size = 1.5, colour = 'blue') +
  geom_line(aes(corr_values, prop2), size = 1.5, colour = 'red') +
  geom_line(aes(corr_values, prop3), size = 1.5, colour = 'green') +
  ylim(c(0, 2)) + 
  labs(x = "Correlation between AQUA - Buoy and VIIRS - Buoy SSTs", y = "Proportion of Actual Variance Predicted") +
  ggsave("Correlation_sim.ps", device = 'ps',
    width = 8, height = 8, units = 'in')

```


```{r correcting_glob_experi_for_corr}

corrections <- c(prop1[71], prop2[71], prop3[71]) # Correlation between aqua and viirs errors ~ 0.7 - get correction values for 0.7
sd_global_corrected <- sd_global / corrections # Correct sds

```

```{r}
mm1 <- orig_filtered$amsr2.val.vrs
mm2 <- which(mm1 > -999.000)
mm1 <- mm1[mm2]
qq1 <- mm1 - orig_filtered$buoy.sst.vrs[mm2]
fit_amsr2_aqua_errors <- lm(qq1 ~ tt1[mm2])
fit_amsr2_viirs_errors <- lm(qq1 ~ tt2[mm2])

```











