---
title: "Triple Collocation of VIIRS, AMSR-2, and Buoy Matchups"
output:
  html_notebook:
    toc: True
    theme: united
---


This notebook describes the steps necessary to conduct a triple collocation (TCL)
study of SST values derived from (a) Suomi_NPP - VIIRS, (b) GCOM-W - AMSR-2 and
(c) in situ observations form moored and drifting buoys.

```{r prep_workspace, include=FALSE}
# Load necessary packages
library(dplyr)
library(ggplot2)
library(mapview)
library(futile.logger)
library(lubridate)
library(tidyr)
require(RColorBrewer)

ggplot <- function(...) {ggplot2::ggplot(...) + theme_bw()}

# Source own functions
# Define direwctory where functions are for each operating system

if (Sys.info()["sysname"] == 'Windows') {
  fun.dir <- 'D:/matchups/r-projects/Matchup_R_Scripts/Functions/'
} else if (Sys.info()["sysname"] == 'Linux') {
  fun.dir <- '/home/ckk/Projects/Matchup_R_Scripts/Functions/'
}

fun.file <- paste0(fun.dir, 'common_functions.R')

source(file = fun.file,
  local = FALSE, echo = FALSE, verbose = FALSE)
rm(fun.dir, fun.file)
```


# Building a Dataset for Triple Collocation

## Loading Matchups

In a previous step, we read in matchups VIIRS and AMSR2 and stored them in
binary (.Rdata) objects. The AMSR2 retrievals are matched up with VIIRS matchups
and stored in the same binary object. The first step is to restore that object. Note that
the name of the stored binary objects (.Rdata) contain the date in which they
were created. Be sure to edit the names of the objects to ensure that the most
recent versions are being used.  

```{r load_VIIRS, eval=FALSE, include=FALSE}

# Load .Rdata object for VIIRS matchups resulting from script 01_read_matchups.R
# We assume... TO DO

if (Sys.info()["sysname"] == 'Windows') {
  # Windows directory for VIIRS matchups
  win.dir <- 'D:/matchups/r-projects/R_MUDB/viirs/v641ao/mia/results/objects/'
  # Name of .Rdata file (update as needed)
  win.file <- 'VIIRS_Suomi_NPP_MIA_L2GEN_ALL_Class_6.4.1_ao_2016_11_23_with_ancillary.Rdata'
  viirs.file <- paste0(win.dir, win.file)
  if (!file.exists(viirs.file)) {
    stop('Input file does not exist')
  } else {
    load(viirs.file, verbose = TRUE)
  }
  rm(win.dir, win.file, viirs.file)
} else if (Sys.info()["sysname"] == 'Linux') {
  # Linux directory for VIIRS matchups
  # For 787
  linux.dir <- '~/Projects/Matchup_R_Scripts/Results/objects/'
  # For laptop
  #linux.dir <- '~/Projects/Matchup_R_Scripts/'
  # Name of .Rdata file (update as needed)
  linux.file <- 'VIIRS_Suomi_NPP_MIA_L2GEN_ALL_Class_6.4.1_ao_2016_11_22_with_ancillary.Rdata'
  viirs.file <- paste0(linux.dir, linux.file)
  if (!file.exists(viirs.file)) {
    stop('Input file does not exist')
  } else {
    load(viirs.file, verbose = TRUE)
  }
  rm(linux.dir, linux.file, viirs.file)
}

# Create 'VIIRS' object so it has a shorter name
VIIRS <- VIIRS_Suomi_NPP_MIA_L2GEN_ALL_Class_6.4.1_ao_2016_11_22
rm(VIIRS_Suomi_NPP_MIA_L2GEN_ALL_Class_6.4.1_ao_2016_11_22)
```

The restored VIIRS matchups span the period between `r min(VIIRS$buoy.timedate)` 
and `r max(VIIRS$buoy.timedate)`. The original VIIRS dataset had `r nrow(VIIRS)` matchups.
The AMSR2 data we have is simply an SST value that matched-up to a VIIRS matchup.
We cannot compute much of the same info for AMSR2 matchups but know we have 
`r length(VIIRS$amsr2.val[VIIRS$amsr2.val != -999])` AMSR2 retrievals.

## Prepping Matchups

The VIIRS object contains many more variables than actually needed for our TCL analysis.
We start with `r ncol(VIIRS)` variables and cut many that are not necessary.
The VIIRS object contains both AMSR2 and VIIRS SSTs. Both of these SSTs must be prepped 
for our TCL analysis. First, we take AMSR2 values that are -999 and turn them into NAs. 
We then convert VIIRS SSTs to Celsius from Kelvin. AMSR2 SSTs are already in Celsius.
Finally, we convert VIIRS SSTs to bulk SSTs from skin SSTs. AMSR2 SSTs are already bulk SSTs.

```{r prep_VIIRS, include=FALSE}
# AMSR-2 data does not come with a timestamp
# We are interested in exploring the correlation between AMSR and VIIRS data
# These are some preliminary steps to begin that process

# Not all variables in the VIIRS dataframes are necessary for our
# error analysis
# We cut some variables here
# AMSR-E was put to sleep in 2011 and there is no overlap between
# AMSR-E and VIIRS
# To elminate confusion, the AMSR-E column is cut here

viirs.colloc.vars <- keep_matchup_variables_collocation(matchup.format = 'MIA_L2GEN',
  sensor = 'VIIRS', ancillary.data = TRUE)

VIIRS <- dplyr::select(VIIRS, one_of(viirs.colloc.vars))

# Some AMSR-2/WSAT data has a value of -999 when it should be NA
# We convert those here
# Get an idea of how many matchups this is...
table(VIIRS$amsr2.val == -999)

VIIRS$amsr2.val <- ifelse(VIIRS$amsr2.val == -999, NA, VIIRS$amsr2.val)

# VIIRS data is in Kelvin while AMSR-2 is in Celsius
# We convert VIIRS' Kelvin to Celsius here

pattern <- paste("^cen", "^med", "^max", "^min", sep = "|")
tt1 <- which((grepl(pattern, names(VIIRS))))

# Convert selected columns from K to C
VIIRS[tt1] <- apply(VIIRS[tt1], 2, FUN = degK.to.degC)

# Finally, VIIRS and AMSR-2 gather different types of SST
# VIIRS is a skin measurement penetrating just a few millimeteres
# AMSR-2 penetrates a few microns
# We apply a correction here
# VIIRS skin-to-buoy (bulk SST)

VIIRS$cen.sst <- VIIRS$cen.sst + 0.17

# VIIRS sat and buoy dates are characters
# Convert them to POSIXct for ease

VIIRS$buoy.date <- lubridate::ymd(VIIRS$buoy.date)
VIIRS$sat.date <- lubridate::ymd(VIIRS$sat.date)
```

We now have `r ncol(VIIRS)` variables.

## Filtering Matchups

We now filter the VIIRS object so we have only good quality, nighttime, buoy SST retrievals. 
This filtering is only done for VIIRS matchups as for AMSR2 retrievals we do not have have this info.
However, because of the matching up process between VIIRS and AMSR2 retrievals, VIIRS buoy retrievals
will probably always be AMSR2 buoy retrievals. Nighttime retrievals for VIIRS are probably also nighttime
retrievals for AMSR2. However, nighttime does not impact AMSR2 retrieval accuracy like it impacts VIIRS
retrieval accuracy. We do not have quality levels for AMSR2 retrievals so no quality filtering can
be done for AMSR2 retrievals. However, different geophyisical conditions impact retrieval accuracy
for AMSR2 retrievals than VIIRS retrievals. Finally, we cannot use a retrieval if either VIIRS or AMSR2
have an NA as the SST entry. We filter out all such retrievals.

```{r filter_viirs, echo=TRUE}
VIIRS <- dplyr::tbl_df(VIIRS) %>%
  dplyr::filter(qsst == 0) %>%
  dplyr::filter(solz >= 90) %>%
  dplyr::filter(insitu.platform %in% c("DriftingBuoy","MooredBuoy")) %>%
  dplyr::filter(!is.na(cen.sst)) %>%
  dplyr::filter(!is.na(amsr2.val))
```

We now have `r nrow(VIIRS)` VIIRS matchups and `r length(VIIRS$amsr2.val[VIIRS$amsr2.val != NA])` AMSR2 matchups.
We have collocated AMSR2 matchups which are attached to our VIIRS matchups (the matchups are processed like this)
and, because of the filtering we have applied, the VIIRS matchups are now ready for a TCL analysis.

# Exploring the TC Dataset

Before doing the TCL analysis, we explore the dataset we have generated.


## Spatial Distribution of Matchups

Here we explore the spatial distribution of collocated matchups.

```{r spatial_distribution, echo=TRUE}
# Create base map
require(ggmap)

mp <- NULL
mapWorld <- borders("world", colour = "gray50", fill = "gray70") # create a layer of borders
mp <- ggplot() +  mapWorld +
  ggplot2::scale_x_continuous(breaks = seq(from = -180, to = 180, by = 60)) +
  ggplot2::scale_y_continuous(breaks = seq(from = -90, to = 90, by = 30)) +
  coord_fixed(ratio = 1)

# WARNING: Be careful as this worked with ggplot2 2.1.0 on Linux
# Try to work on different machines with varying versions of ggplot2
# Now layer the buoys on top with hexagonal binning
mp <- mp + 
  ggplot2::stat_binhex(data = VIIRS,
    aes(x = buoy.lon, # Use hexagonal binning command and give x and y input
    y = buoy.lat,
    fill = cut(..count.., c(0, 100, 500, 1000, 1500, Inf))), # Divides matchups per bin into discrete chunks and colors likewise
    binwidth = c(10, 10)) +
  mapWorld + labs(x = NULL, y = NULL) +
  #scale_fill_hue('value') + # Standard colors with discrete chunking
  scale_fill_brewer(palette = 'YlOrRd') + # Change colors to Yellow, Orange, and Red - many diff 
  guides(fill = guide_legend(title = "N of matchups"))

#ggplot2::ggsave(filename = 'spatial_distribution.ps', device = 'ps',
#       width = 8, height = 6, units = 'in')

mp
```

## Temporal Distribution of Matchups

```{r temporal_distribution}

yy <- lubridate::year(VIIRS$buoy.timedate)     # Year
mm <- lubridate::month(VIIRS$buoy.timedate)    # Month

tt1 <- cbind(yy, mm)

tt2 <- dplyr::tbl_df(tt1) %>%
  dplyr::group_by(yy, mm) %>%
  dplyr::summarise(n = n()) %>%
  dplyr::arrange(yy, mm)

yy.vals <- sort(unique(yy))
mm.vals <- sort(unique(mm))
tt3 <- expand.grid(mm = mm.vals, yy = yy.vals)

tt4 <- dplyr::left_join(tt3, tt2) %>%
  dplyr::select(yy, mm, n)

brks <- pretty(tt4$n, n = 5)
#brks <- floor(quantile(tt4$n, probs = seq(0, 1, 0.25), na.rm = TRUE))

tt4$n <- cut(tt4$n, breaks = brks)
tt4$mm <- ordered(tt4$mm, labels = month.abb)
tt4$yy <- ordered(tt4$yy)

hm <- ggplot2::ggplot(data = tt4, aes(mm, yy)) +
  ggplot2::geom_tile(aes(fill = n), colour = "white") +
  ggplot2::scale_fill_brewer(type = "seq", palette = 'YlOrRd', direction = 1) +
  ggplot2::theme_bw() +
  ggplot2::theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  #ggplot2::ggsave("Temporal_heatmap.ps", device = 'ps',
       #width = 8, height = 6, units = 'in')
```

## Association Between all SST Sources

Here we explore the correlation between SST sources in a series of pairwise plots.
The nature of systematic errors in one source can be seen from these plots

```{r viirs_buoy_corr, echo=TRUE}
# VIIRS and buoy correlation
ccc <- RColorBrewer::brewer.pal(n = 8, name = 'YlOrRd')

ggplot(data = VIIRS, mapping = aes(x = buoy.sst, y = cen.sst)) +
  stat_bin2d(binwidth = c(0.2, 0.1)) +
  scale_fill_gradientn(colours = ccc) +
  geom_abline(intercept = 0, slope = 1, color = 'steelblue') +
  lims(x = c(-2, 32), y = c(-2,32)) +
  ggplot2::labs(title = "VIIRS and Buoy SSTs", x = 'Buoy SST', y = 'VIIRS SST') +
  #ggplot2::ggsave(filename = 'VIIRS_and_buoy_SSTs.ps', device = 'ps',
       #width = 8, height = 6, units = 'in') +
  theme(text = element_text(size = 16))

fit_viirs_buoy_sst <- lm(buoy.sst ~ cen.sst, data = VIIRS)

# Compute robust standard deviation

RSD_viirs_buoy_sst <- IQR(fit_viirs_buoy_sst$residuals) / 1.348
```


```{r amsr2_buoy_corr, echo=TRUE}
# AMSR-2 and buoy correlation
ccc <- RColorBrewer::brewer.pal(n = 8, name = 'YlOrRd')

ggplot(data = VIIRS, mapping = aes(x = buoy.sst, y = amsr2.val)) +
  stat_bin2d(binwidth = c(0.2, 0.1)) +
  scale_fill_gradientn(colours = ccc) +
  geom_abline(intercept = 0, slope = 1, color = 'steelblue') +
  lims(x = c(-2, 32), y = c(-2,32)) +
  ggplot2::labs(title = "AMSR2 and Buoy SSTs", x = 'Buoy SST', y = 'AMSR-2 SST') +
  #ggplot2::ggsave(filename = 'VIIRS_and_AMSR2_SSTs.ps', device = 'ps',
       #width = 8, height = 6, units = 'in') +
  theme(text = element_text(size = 16))

fit_amsr2_buoy_sst <- lm(buoy.sst ~ amsr2.val, data = VIIRS)

# Compute robust standard deviation

RSD_amsr2_buoy_sst <- IQR(fit_amsr2_buoy_sst$residuals) / 1.348
```


```{r correlation_between_VIIRS_AMSR-2, echo=TRUE}
# Despite not knowing the exact timestamp of the AMSR-2 data, 
# we can still explore the correlation between 'matched-up'
# VIIRS and AMSR-2 data

ccc <- RColorBrewer::brewer.pal(n = 8, name = 'YlOrRd')

ggplot(data = VIIRS, mapping = aes(x = cen.sst, y = amsr2.val)) +
  stat_bin2d(binwidth = c(0.2, 0.1)) +
  scale_fill_gradientn(colours = ccc) +
  geom_abline(intercept = 0, slope = 1, color = 'steelblue') +
  lims(x = c(-2, 32), y = c(-2,32)) +
  ggplot2::labs(title = "VIIRS and AMSR2 SSTs", x = 'VIIRS SST', y = 'AMSR-2 SST') +
  #ggplot2::ggsave(filename = 'VIIRS_and_AMSR2_SSTs.ps', device = 'ps',
       #width = 8, height = 6, units = 'in') +
  theme(text = element_text(size = 16))

fit_viirs_amsr2_sst <- lm(cen.sst ~ amsr2.val, data = VIIRS)

# Compute robust standard deviation

RSD_viirs_amsr2_sst <- IQR(fit_viirs_amsr2_sst$residuals) / 1.348
```

A key assumption in the TCL analysis is that the random errors across sources are uncorrelated. 
We can estimate the random error by assuming the buoy SST is ground truth and subtracting it from
both the VIIRS and AMSR2 SSTs. We then plot the 'errors' against each other and observe their
correlation.

```{r VIIRS_AMSR2_errors, echo=TRUE}

# By assuming the buoy data is correct, we can estimate
# the error in VIIRS and AMSR-2 data by subtracting the
# buoy SST we have from the VIIRS and AMSR-2 SSTs
ccc <- RColorBrewer::brewer.pal(n = 8, name = 'YlOrRd')
ddd <- VIIRS$cen.sst - VIIRS$buoy.sst
eee <- VIIRS$amsr2.val - VIIRS$buoy.sst

ggplot(data = VIIRS, mapping = aes(x = ddd, y = eee)) +
  stat_bin2d(binwidth = c(0.2, 0.1)) +
  scale_fill_gradientn(colours = ccc) +
  geom_abline(intercept = 0, slope = 0, color = 'steelblue') +
  lims(x = c(-2.5, 2.5), y = c(-2.5, 2.5)) +
  ggplot2::labs(title = "VIIRS and AMSR-2 Errors", x = 'VIIRS Errors', y = 'AMSR-2 Errors') +
  #ggplot2::ggsave(filename = 'VIIRS_and_AMSR2_errors.ps', device = 'ps',
       #width = 8, height = 6, units = 'in') +
  theme(text = element_text(size = 16))

fit_viirs_amsr2_errors <- lm(ddd ~ eee)

# Compute robust standard deviation

RSD_viirs_amsr2_errors <- IQR(fit_viirs_amsr2_errors$residuals) / 1.348
```

The correlation among VIIRS and AMSR2 errors is minimal based on our above plot.

# TCL Analysis

# We now use the TCL equations to estimate the standard deviation of therandom error in 
the global TCL dataset.

```{r collocation_functions, include=FALSE}

# Bias between sources
compute_pairwise_bias <- function(src1, src2, src3) {
  if (!is.null(nrow(src1)) |
    !is.null(nrow(src2)) |
    !is.null(nrow(src3)) |
    length(src1) <= 1 |
    length(src2) <= 1 |
    length(src3) <= 1) {
    stop("Only vectors can be passed as arguments to the calculate_bias function.")
  }

  bias_src1src3 <- mean(src1) - mean(src3) #Bias between source1 and source2
  
  bias_src2src3 <- mean(src2) - mean(src3) #Bias between source2 and source3
  
  # Return a dataframe of biases that can re-scale each observation type
  
  biases <- c(bias_src1src3, bias_src2src3)
  names(biases) <- c("bias_src1src3", "bias_src2src3")
  
  return(biases)
}   # End of compute_bias

# Estimates of variance of random error
compute_variance <- function(src1, src2, src3) {
  if (!is.null(nrow(src1)) |
    !is.null(nrow(src2)) |
    !is.null(nrow(src3)) |
    length(src1) <= 1 |
    length(src2) <= 1 |
    length(src3) <= 1) {
    stop("Only vectors can be passed as arguments to the calculate_variance function")
  }
  
  # Call bias function to re-scale measurements
  biases <- compute_pairwise_bias(src1, src2, src3)
  
  #Scale sources to one source to compute variance
  src1_scaled <- src1 - biases['bias_src1src3']
  src2_scaled <- src2 - biases['bias_src2src3']
  
  # Variance of errors in observation types
  var_src1 <- mean((src1_scaled - src2_scaled) * (src1_scaled - src3))              # Variance of source1
  var_src2 <- mean((src1_scaled - src2_scaled) * (src2_scaled - src3))              # Variance of source2
  var_src3 <- mean((src1_scaled - src3) * (src2_scaled - src3))                     # Variance of source3
  
  # Take all the individual variances and return them as a dataframe
  variances <- data.frame(var_src1, var_src2, var_src3)
  colnames(variances) <- c("var_src1", "var_src2", "var_src3")
  return(variances)
} # end of compute_variance

```

```{r run_collocation_global}

# Make sure the last argument to compute_variance corresponds to the buoy SSTs

variances_global <- compute_variance(VIIRS$cen.sst,
  VIIRS$amsr2.val,
  VIIRS$buoy.sst)

biases_global <- compute_pairwise_bias(VIIRS$cen.sst,
  VIIRS$amsr2.val,
  VIIRS$buoy.sst)

sd_global <- round(sqrt(abs(variances_global)), 3)

sd_global

```