---
title: "Triple Collocation of VIIRS, AMSR-2, and Buoy Matchups"
output:
  html_notebook:
    toc: True
    theme: united
---


This notebook describes the steps necessary to conduct a triple collocation
study of SST values derived from (a) Suomi_NPP - VIIRS, (b) GCOM-W - AMSR-2 and
(c) in situ observations form moored and drifting buoys.

```{r prep_workspace, include=FALSE}
# Load necessary packages
library(dplyr)
library(ggplot2)
#library(mapview)
library(futile.logger)
library(lubridate)
library(tidyr)
require(RColorBrewer)

ggplot <- function(...) {ggplot2::ggplot(...) + theme_bw()}

# Source own functions
# Define direwctory where functions are for each operating system

if (Sys.info()["sysname"] == 'Windows') {
  fun.dir <- 'D:/matchups/r-projects/Matchup_R_Scripts/Functions/'
} else if (Sys.info()["sysname"] == 'Linux') {
  fun.dir <- '/home/ckk/Projects/Matchup_R_Scripts/Functions/'
}

fun.file <- paste0(fun.dir, 'common_functions.R')

source(file = fun.file,
  local = FALSE, echo = FALSE, verbose = FALSE)
rm(fun.dir, fun.file)
```


# Building a Dataset for Triple Collocation

In a previous step, we read in matchups for Aqua and VIIRS and stored them in
binary (.Rdata) objects. For Aqua, we modified the regular expression to list 
files so that only matchups from 2010 onwards were read. The first step is to
restore those objects. Note that the name of the stored binary objects (.Rdata)
contain the date in which they were created. Be sure to edit the names of the
objects to ensure that the most recent versions are being used.  

```{r load_VIIRS, eval=False, include=FALSE}

# Load .Rdata object for VIIRS matchups resulting from script 01_read_matchups.R
# We assume... TO DO

if (Sys.info()["sysname"] == 'Windows') {
  # Windows directory for VIIRS matchups
  win.dir <- 'D:/matchups/r-projects/R_MUDB/viirs/v641ao/mia/results/objects/'
  # Name of .Rdata file (update as needed)
  win.file <- 'VIIRS_Suomi_NPP_MIA_L2GEN_ALL_Class_6.4.1_ao_2016_11_23_with_ancillary.Rdata'
  viirs.file <- paste0(win.dir, win.file)
  if (!file.exists(viirs.file)) {
    stop('Input file does not exist')
  } else {
    load(viirs.file, verbose = TRUE)
  }
  rm(win.dir, win.file, viirs.file)
} else if (Sys.info()["sysname"] == 'Linux') {
  # Linux directory for VIIRS matchups
  # For 787
  #linux.dir <- '~/Projects/Matchup_R_Scripts/Results/objects/'
  # For laptop
  linux.dir <- '~/Projects/Matchup_R_Scripts/'
  # Name of .Rdata file (update as needed)
  linux.file <- 'VIIRS_Suomi_NPP_MIA_L2GEN_ALL_Class_6.4.1_ao_2016_11_22_with_ancillary.Rdata'
  viirs.file <- paste0(linux.dir, linux.file)
  if (!file.exists(viirs.file)) {
    stop('Input file does not exist')
  } else {
    load(viirs.file, verbose = TRUE)
  }
  rm(linux.dir, linux.file, viirs.file)
}

# Create 'VIIRS' object so it has a shorter name
VIIRS <- VIIRS_Suomi_NPP_MIA_L2GEN_ALL_Class_6.4.1_ao_2016_11_22
rm(VIIRS_Suomi_NPP_MIA_L2GEN_ALL_Class_6.4.1_ao_2016_11_22)
```

```{r load_AQUA, eval=FALSE, include=FALSE}
# Load .Rdata object for AQUA matchups resulting from script 01_read_matchups.R

if (Sys.info()["sysname"] == 'Windows') {
  # Windows directory for AQUA matchups
  win.dir <- 'D:/matchups/r-projects/R_MUDB/modis/aqua/gsfc/l6cv6/results/objects/'
  # Name of .Rdata file (update as needed)
  win.file <- 'MODIS_Aqua_GSFC_ALL_Class_6.4.1_ao_2016_11_23_with_ancillary.Rdata'
  viirs.file <- paste0(win.dir, win.file)
  if (!file.exists(viirs.file)) {
    stop('Input file does not exist')
  } else {
    load(viirs.file, verbose = TRUE)
  }
  rm(win.dir, win.file, viirs.file)
} else if (Sys.info()["sysname"] == 'Linux') {
  # Linux directory for VIIRS matchups
  # For 787
  #linux.dir <- '~/Projects/Matchup_R_Scripts/Results/objects/'
  # For laptop
  linux.dir <- '~/Projects/Matchup_R_Scripts/'
  # Name of .Rdata file (update as needed)
  #linux.file <- 'MODIS_Aqua_GSFC_ALL_Class_6.4.1_ao_2016_11_26_with_ancillary.Rdata'
  linux.file <- 'MODIS_Aqua_GSFC_ALL_Class_6.4.1_ao_2017_03_11_with_ancillary.Rdata'
  aqua.file <- paste0(linux.dir, linux.file)
  if (!file.exists(aqua.file)) {
    stop('Input file does not exist')
  } else {
    load(aqua.file, verbose = TRUE)
  }
  rm(linux.dir, linux.file, aqua.file)
}

# Create 'AQUA' object so it has a shorter name
AQUA <- MODIS_Aqua_GSFC_ALL_Class_6.4.1_ao_2017_03_11
rm(`MODIS_Aqua_GSFC_ALL_Class_6.4.1_ao_2017_03_11`)
```

```{r prep_VIIRS, include=FALSE}
# AMSR-2 data does not come with a timestamp
# We are interested in exploring the correlation between AMSR and VIIRS data
# These are some preliminary steps to begin that process

# Some AMSR-2/WSAT data has a value of -999 when it should be NA
# We convert those here
# Get an idea of how many matchups this is...
table(VIIRS$amsr2.val == -999)

VIIRS$amsr2.val <- ifelse(VIIRS$amsr2.val == -999, NA, VIIRS$amsr2.val)

# Not all variables in the VIIRS dataframes are necessary for our
# error analysis
# We cut some variables here
# AMSR-E was put to sleep in 2011 and there is no overlap between
# AMSR-E and VIIRS
# To elminate confusion, the AMSR-E column is cut here

viirs.colloc.vars <- keep_matchup_variables_collocation(matchup.format = 'MIA_L2GEN',
  sensor = 'VIIRS', ancillary.data = TRUE)

VIIRS <- dplyr::select(VIIRS, one_of(viirs.colloc.vars))

# VIIRS data is in Kelvin while AMSR-2 is in Celsius
# We convert VIIRS' Kelvin to Celsius here

pattern <- paste("^cen", "^med", "^max", "^min", sep = "|")
tt1 <- which((grepl(pattern, names(VIIRS))))

# Convert selected columns from K to C
VIIRS[tt1] <- apply(VIIRS[tt1], 2, FUN = degK.to.degC)

# Finally, VIIRS and AMSR-2 gather different types of SST
# VIIRS is a skin measurement penetrating just a few millimeteres
# AMSR-2 penetrates a few microns
# We apply a correction here
# VIIRS skin-to-buoy (bulk SST)

VIIRS$cen.sst <- VIIRS$cen.sst + 0.17
```


```{r viirs_buoy_corr, echo=TRUE}
# VIIRS and buoy correlation
ccc <- RColorBrewer::brewer.pal(n = 8, name = 'YlOrRd')

ggplot(data = VIIRS, mapping = aes(x = buoy.sst, y = cen.sst)) +
  stat_bin2d(binwidth = c(0.2, 0.1)) +
  scale_fill_gradientn(colours = ccc) +
  geom_abline(intercept = 0, slope = 1, color = 'steelblue') +
  lims(x = c(-2, 32), y = c(-2,32)) +
  ggplot2::labs(title = "VIIRS and Buoy SSTs", x = 'Buoy SST', y = 'VIIRS SST') +
  ggplot2::ggsave(filename = 'VIIRS_and_buoy_SSTs.ps', device = 'ps',
       width = 8, height = 6, units = 'in') +
  theme(text = element_text(size = 16))

fit_viirs_buoy_sst <- lm(buoy.sst ~ cen.sst, data = VIIRS)

# Compute robust standard deviation

RSD_viirs_buoy_sst <- IQR(fit_viirs_buoy_sst$residuals) / 1.348
```


```{r amsr-2_buoy_corr, echo=TRUE}
# AMSR-2 and buoy correlation
ccc <- RColorBrewer::brewer.pal(n = 8, name = 'YlOrRd')

ggplot(data = VIIRS, mapping = aes(x = buoy.sst, y = amsr2.val)) +
  stat_bin2d(binwidth = c(0.2, 0.1)) +
  scale_fill_gradientn(colours = ccc) +
  geom_abline(intercept = 0, slope = 1, color = 'steelblue') +
  lims(x = c(-2, 32), y = c(-2,32)) +
  ggplot2::labs(title = "AMSR-2 and Buoy SSTs", x = 'Buoy SST', y = 'AMSR-2 SST') +
  ggplot2::ggsave(filename = 'VIIRS_and_AMSR-2_SSTs.ps', device = 'ps',
       width = 8, height = 6, units = 'in') +
  theme(text = element_text(size = 16))

fit_amsr2_buoy_sst <- lm(buoy.sst ~ amsr2.val, data = VIIRS)

# Compute robust standard deviation

RSD_amsr2_buoy_sst <- IQR(fit_amsr2_buoy_sst$residuals) / 1.348
```


```{r correlation_between_VIIRS_AMSR-2, echo=TRUE}
# Despite not knowing the exact timestamp of the AMSR-2 data, 
# we can still explore the correlation between 'matched-up'
# VIIRS and AMSR-2 data

ccc <- RColorBrewer::brewer.pal(n = 8, name = 'YlOrRd')

ggplot(data = VIIRS, mapping = aes(x = cen.sst, y = amsr2.val)) +
  stat_bin2d(binwidth = c(0.2, 0.1)) +
  scale_fill_gradientn(colours = ccc) +
  geom_abline(intercept = 0, slope = 1, color = 'steelblue') +
  lims(x = c(-2, 32), y = c(-2,32)) +
  ggplot2::labs(title = "VIIRS and AMSR-2 SSTs", x = 'VIIRS SST', y = 'AMSR-2 SST') +
  ggplot2::ggsave(filename = 'VIIRS_and_AMSR-2_SSTs.ps', device = 'ps',
       width = 8, height = 6, units = 'in') +
  theme(text = element_text(size = 16))

fit_viirs_amsr2_sst <- lm(cen.sst ~ amsr2.val, data = VIIRS)

# Compute robust standard deviation

RSD_viirs_amsr2_sst <- IQR(fit_viirs_amsr2_sst$residuals) / 1.348
```

```{r VIIRS_AMSR2_errors, echo=TRUE}

# By assuming the buoy data is correct, we can estimate
# the error in VIIRS and AMSR-2 data by subtracting the
# buoy SST we have from the VIIRS and AMSR-2 SSTs
ccc <- RColorBrewer::brewer.pal(n = 8, name = 'YlOrRd')
ddd <- VIIRS$cen.sst - VIIRS$buoy.sst
eee <- VIIRS$amsr2.val - VIIRS$buoy.sst

ggplot(data = VIIRS, mapping = aes(x = ddd, y = eee)) +
  stat_bin2d(binwidth = c(0.2, 0.1)) +
  scale_fill_gradientn(colours = ccc) +
  geom_abline(intercept = 0, slope = 1, color = 'steelblue') +
  lims(x = c(-10, 10), y = c(-10, 10)) +
  ggplot2::labs(title = "VIIRS and AMSR-2 Errors", x = 'VIIRS Errors', y = 'AMSR-2 Errors') +
  ggplot2::ggsave(filename = 'VIIRS_and_AMSR2_errors.ps', device = 'ps',
       width = 8, height = 6, units = 'in') +
  theme(text = element_text(size = 16))

fit_viirs_amsr2_errors <- lm(ddd ~ eee)

# Compute robust standard deviation

RSD_viirs_amsr2_errors <- IQR(fit_viirs_amsr2_errors$residuals) / 1.348
```