---
title: "Exploration of a MODIS SSES Dataset - Algorithms"
author: "Chirag Kumar and Guillermo Podesta"
output:
  html_notebook:
    toc: True
    theme: united
---

# Prep Workspace and Data
We use MODIS matchups from 2002 - 2016 that have been read in through a previous script. Here we simply load
the matchups, any necessary packages, and define necessary functions.

```{r prep_workspace, include=FALSE}
# Import necessary packages
require(ggplot2)
require(dplyr)
require(RColorBrewer)
require(circular)
require(ggmap)
require(raster)
require(sfsmisc)
require(rasterVis)
require(rgdal)

ggplot <- function(...) {ggplot2::ggplot(...) + theme_bw()}

# Define secant function
secant.deg <- function(x) {1 / (cos(circular::rad(x)))}

# Source own functions
# Define direwctory where functions are for each operating system

if (Sys.info()["sysname"] == 'Windows') {
  fun.dir <- 'D:/matchups/r-projects/Matchup_R_Scripts/Functions/'
} else if (Sys.info()["sysname"] == 'Linux') {
  fun.dir <- '/home/ckk/Projects/Matchup_R_Scripts/Functions/'
}

fun.file <- paste0(fun.dir, 'common_functions.R')

source(file = fun.file,
  local = FALSE, echo = FALSE, verbose = FALSE)
rm(fun.dir, fun.file)


# Load data
# For 787
linux_dir <- '~/Projects/Matchup_R_Scripts/Results/objects/'
# For Laptop
#linux_dir <- '~/Projects/Matchup_R_Scripts/'
linux_file <- 'MODIS_Aqua_GSFC_ALL_Class_6.4.1_ao_2017_04_12_with_ancillary.Rdata'
AQUA_file <- paste0(linux_dir, linux_file)

if (!file.exists(AQUA_file)) {
    stop('Input file does not exist')
  } else {
    load(AQUA_file, verbose = TRUE)
    AQUA <- MODIS_Aqua_GSFC_ALL_Class_6.4.1_ao_2017_04_12
  }

rm(linux_dir, linux_file, AQUA_file)
rm(MODIS_Aqua_GSFC_ALL_Class_6.4.1_ao_2017_04_12)
```

The matchup file read in at a previous step is not filtered and contains too many variables. For the purpose of 
this analysis, we use only nighttime matchups with a quality of 0, 1, or 2 and only keep variables from the infrared
channels we use.

```{r prep_data, echo=TRUE}
# Turn POSIXct objects to characters bc dplyr doesn't support POSIXct
AQUA$sat.timedate <- as.character(AQUA$sat.timedate)
AQUA$buoy.timedate <- as.character(AQUA$buoy.timedate)

# Apply basic filtering
AQUA <- dplyr::tbl_df(AQUA) %>%
  dplyr::filter(solz >= 90) %>%
  dplyr::filter(qsst == 0 | qsst == 1 | qsst == 2)

# Now grab only variables that may be used to determine retrieval accuracy and make some new variables (i.e. x1, x2, x3)
# Create df of features
orig <- dplyr::tbl_df(AQUA) %>%
  dplyr::mutate(x1 = cen.11000,
    x2 = cen.11000 - cen.12000,
    x3 = cen.ref.type.1.SST,
    satz = satz,
    x2 = x2 * x3,
    x3 = ((secant.deg(satz) - 1) * x2),
    lat = buoy.lat,
    lon = buoy.lon,
    sd11 = sd.11000,
    sd12 = sd.12000,
    range11 = max.11000 - min.11000,
    range12 = max.12000 - min.12000,
    diff.med.min11 = med.11000 - min.11000,
    diff.med.min12 = med.12000 - min.12000,
    band.diff = cen.11000 - cen.12000,
    qsst = qsst,
    buoy.sst = buoy.sst,
    cell5deg = cell5deg,
    buoy.timedate = buoy.timedate,
    SST.resid.SMB = cen.sst - buoy.sst) %>% # SMB = sat minus buoy - also don't debias SSTs i.e. turn buoy into skin
  dplyr::select(x1, x2, x3, satz, lon, lat, sd11, sd12, range11, range12, diff.med.min11, diff.med.min12, band.diff, qsst, buoy.sst, cell5deg,
                buoy.timedate, SST.resid.SMB)

```


# Relationship Between band_diff and temp_deficit

```{r band_diff_temp_deficit_prediction_buoy_sst}

zz1 <- dplyr::tbl_df(orig) %>%
  dplyr::filter(qsst == 0)

ccc = RColorBrewer::brewer.pal(n = 8, 'YlOrRd')
ccc <- ccc[3:8]

temp_context <- cut(zz1$buoy.sst, c(-Inf, 10, 25, Inf))
temp_deficit <- zz1$buoy.sst - zz1$x1

uuu <- data.frame(temp_context = as.factor(temp_context), temp_deficit = temp_deficit, band.diff = zz1$band.diff)

# Take a sample of uuu bc loess requires lots of memory - for testing purposes
#set.seed(42)
#index <- sample(1:nrow(uuu), 1000, replace = FALSE)
#uuu <- uuu[index, ]

ggplot2::ggplot(data = uuu, aes(x = temp_deficit, y = band.diff)) +
  geom_bin2d(bins = 100) +
  scale_fill_gradientn(colours = ccc) + 
  geom_smooth(method = loess, se = FALSE, span = 0.5) +
  facet_wrap(~temp_context, nrow = 3, ncol = 1) +
  ggsave('band_diff_predictor_of_temp_deficit_different_in_situ_windows.pdf')

# Regression equations and statistics for the three facets
uuu$temp_context_numeric <- factor(as.character(uuu$temp_context), labels = c(0, 1, 2), levels = c('(-Inf,10]', '(10,25]', '(25, Inf]'))

temp_less_10_lm <- lm(formula = temp_deficit ~ band.diff, data = uuu[uuu$temp_context_numeric == 0, ])
rSD_temp_less_10 <- IQR(temp_less_10_lm$residuals) / 1.348
cor_temp_less_10 <- cor(uuu[uuu$temp_context_numeric == 0, 'temp_deficit'], uuu[uuu$temp_context_numeric == 0, 'band.diff'])

temp_between_10_25_lm <- lm(formula = temp_deficit ~ band.diff, data = uuu[uuu$temp_context_numeric == 1, ])
rSD_temp_between_10_25 <- IQR(temp_between_10_25_lm$residuals) / 1.348
cor_temp_between_10_25 <- cor(uuu[uuu$temp_context_numeric == 1, 'temp_deficit'], uuu[uuu$temp_context_numeric == 1, 'band.diff'])

temp_greater_25_lm <- lm(formula = temp_deficit ~ band.diff, data = uuu[uuu$temp_context_numeric == 2, ])
rSD_temp_greater_25 <- IQR(temp_greater_25_lm$residuals) / 1.348
cor_temp_greater_25 <- cor(uuu[uuu$temp_context_numeric == 2, 'temp_deficit'], uuu[uuu$temp_context_numeric == 2, 'band.diff'])

# Whole dataset for comparison
temp_all_lm <- lm(formula = temp_deficit ~ band.diff, data = uuu)
rSD_temp_all <- IQR(temp_all_lm$residuals) / 1.348
cor_temp_all <- cor(uuu$temp_deficit, uuu$band.diff)

```































