---
title: "Exploration of a MODIS SSES Dataset - Decision Trees"
author: "Chirag Kumar and Guillermo Podesta"
output:
  html_notebook:
    toc: True
    theme: united
---

# Prep Workspace and Data
We use MODIS matchups from 2002 - 2016 that have been read in through a previous script. Here we simply load
the matchups, any necessary packages, and define necessary functions.

```{r prep_workspace, include=FALSE}
# Import necessary packages
require(ggplot2)
require(dplyr)
require(RColorBrewer)
require(circular)
require(ggmap)
require(raster)
require(sfsmisc)
require(rasterVis)
require(rgdal)
require(rpart)
require(rpart.plot)
require(caret)
require(evtree)

ggplot <- function(...) {ggplot2::ggplot(...) + theme_bw()}

# Define secant function
secant.deg <- function(x) {1 / (cos(circular::rad(x)))}

# Source own functions
# Define direwctory where functions are for each operating system

if (Sys.info()["sysname"] == 'Windows') {
  fun.dir <- 'D:/matchups/r-projects/Matchup_R_Scripts/Functions/'
} else if (Sys.info()["sysname"] == 'Linux') {
  fun.dir <- '/home/ckk/Projects/Matchup_R_Scripts/Functions/'
}

fun.file <- paste0(fun.dir, 'common_functions.R')

source(file = fun.file,
  local = FALSE, echo = FALSE, verbose = FALSE)
rm(fun.dir, fun.file)


# Load data
# For 787
linux_dir <- '~/Projects/Matchup_R_Scripts/Results/objects/'
# For Laptop
#linux_dir <- '~/Projects/Matchup_R_Scripts/'
linux_file <- 'MODIS_Aqua_GSFC_ALL_Class_6.4.1_ao_2017_04_12_with_ancillary.Rdata'
AQUA_file <- paste0(linux_dir, linux_file)

if (!file.exists(AQUA_file)) {
    stop('Input file does not exist')
  } else {
    load(AQUA_file, verbose = TRUE)
    AQUA <- MODIS_Aqua_GSFC_ALL_Class_6.4.1_ao_2017_04_12
  }

rm(linux_dir, linux_file, AQUA_file)
rm(MODIS_Aqua_GSFC_ALL_Class_6.4.1_ao_2017_04_12)
```

The matchup file read in at a previous step is not filtered and contains too many variables. For the purpose of 
this analysis, we use only nighttime matchups with a quality of 0, 1, or 2 and only keep variables from the infrared
channels we use.

```{r prep_data, echo=TRUE}
# Turn POSIXct objects to characters bc dplyr doesn't support POSIXct
AQUA$sat.timedate <- as.character(AQUA$sat.timedate)
AQUA$buoy.timedate <- as.character(AQUA$buoy.timedate)

# Apply basic filtering
AQUA <- dplyr::tbl_df(AQUA) %>%
  dplyr::filter(solz >= 90) %>%
  dplyr::filter(qsst == 0 | qsst == 1 | qsst == 2)

# Now grab only variables that may be used to determine retrieval accuracy and make some new variables (i.e. x1, x2, x3)
# Create df of features
orig <- dplyr::tbl_df(AQUA) %>%
  dplyr::mutate(T_11 = cen.11000,
    x2 = cen.11000 - cen.12000,
    x3 = cen.ref.type.1.SST,
    satz = satz,
    x2 = x2 * x3,
    x3 = ((secant.deg(satz) - 1) * x2),
    lat = buoy.lat,
    lon = buoy.lon,
    sd11 = sd.11000,
    sd12 = sd.12000,
    range11 = max.11000 - min.11000,
    range12 = max.12000 - min.12000,
    diff.med.min11 = med.11000 - min.11000,
    diff.med.min12 = med.12000 - min.12000,
    band.diff = cen.11000 - cen.12000,
    qsst = qsst,
    buoy.sst = buoy.sst,
    cell5deg = cell5deg,
    buoy.timedate = buoy.timedate,
    T_12 = cen.12000,
    ref_SST = cen.ref.type.1.SST,
    SST.resid.SMB = cen.sst - buoy.sst) %>% # SMB = sat minus buoy - also don't debias SSTs i.e. turn buoy into skin
  dplyr::select(T_11, x2, x3, satz, lon, lat, sd11, sd12, range11, range12, diff.med.min11, diff.med.min12, band.diff, qsst, buoy.sst, cell5deg, buoy.timedate, T_12, ref_SST, SST.resid.SMB)

```

# Decision Trees

Here we explore the ability of decision trees to accurately classify SST matchups that fall in defined residual
ranges.  First we prep the data into training and validation sets for the decision tree exploration.

## Prep Data for Decision Trees

```{r prep_data_decision_trees, echo=TRUE}

# Debias SST resid - i.e. account for skin to buil conversion
orig$SST.resid.SMB <- orig$SST.resid.SMB + 0.17

# Make classes
class_resid <- ifelse(orig$SST.resid.SMB > 1, 'resid > 1', ifelse(orig$SST.resid.SMB < -1, 'resid < -1', '-1 <= resid <= 1'))
orig$class_resid <- as.factor(class_resid)

# Divide data into training and validation sets
set.seed(42)
prop_train <- .6
len_train <- floor(prop_train * nrow(orig))
index <- sample(1:nrow(orig), len_train, replace = FALSE)

orig_train <- orig[index, ]
orig_test <- orig[-index, ]

```


## rpart Decision Trees

```{r rpart_decision_tree, echo=TRUE}

# Train a stard rpart decision tree on above data
rpart_tree <- rpart::rpart(formula = class_resid ~ T_11 + T_12 + band.diff + ref_SST + satz + range11 + sd11, 
                           data = orig_train, 
                           minsplit = 100)

# Plot tree
rpart.plot::prp(rpart_tree, uniform = TRUE, faclen = 0, varlen = 0,
  main = "rpart Classification Tree for Identifying Matchups' Residual Range")

# Evaluate the decision tree's accuracy on validation set
rpart_tree_predictions <- predict(rpart_tree, data = orig_test, type = 'class', na.action = na.pass)

# Draw confusion matrix for predictions on rpart tree
rpart_tree_cm <- caret::confusionMatrix(data = rpart_tree_predictions, reference = orig_train$class_resid)
rpart_tree_cm

```


## Evolutionary Tree


```{r evtree_decision_tree, echo=TRUE}

# Train a stard evtree decision tree on above data
evtree_tree <- evtree::evtree(formula = class_resid ~ T_11 + T_12 + band.diff + ref_SST + satz + range11 + sd11,
                              data = orig_train, 
                              minsplit = 100)

# Plot tree
plot(evtree_tree)

# Evaluate the decision tree's accuracy on validation set
evtree_tree_predictions <- predict(evtree_tree, data = orig_test, type = 'class', na.action = na.pass)

# Draw confusion matrix for predictions on evtree tree
evtree_tree_cm <- caret::confusionMatrix(data = evtree_tree_predictions, reference = orig_train$class_resid)
evtree_tree_cm

```










