---
title: "Exploration of a MODIS SSES Dataset - NLSST Justification AGU"
author: "Chirag Kumar and Guillermo Podesta"
output:
  html_notebook:
    toc: True
    theme: united
---


# Prep Workspace and Data
We use MODIS matchups from 2002 - 2016 that have been read in through a previous script. Here we simply load
the matchups, any necessary packages, and define necessary functions.

```{r prep_workspace, include=FALSE}
# Import necessary packages
require(ggplot2)
require(dplyr)
require(RColorBrewer)
require(circular)
require(ggmap)
require(raster)
require(sfsmisc)
require(rasterVis)
require(rgdal)
require(rpart)
require(rpart.plot)
require(caret)
require(evtree)
require(ctree)
require(partykit)
require(randomForest)
require(RWeka)
require(DMwR)
require(wesanderson)

ggplot <- function(...) {ggplot2::ggplot(...) + theme_bw()}

# Define secant function
secant.deg <- function(x) {1 / (cos(circular::rad(x)))}

# Define function to turn output of gettree from random forest into a dendrogram
to.dendrogram <- function(dfrep,rownum=1,height.increment=0.1){

  if(dfrep[rownum,'status'] == -1){
    rval <- list()

    attr(rval,"members") <- 1
    attr(rval,"height") <- 0.0
    attr(rval,"label") <- dfrep[rownum,'prediction']
    attr(rval,"leaf") <- TRUE

  }else{##note the change "to.dendrogram" and not "to.dendogram"
    left <- to.dendrogram(dfrep,dfrep[rownum,'left daughter'],height.increment)
    right <- to.dendrogram(dfrep,dfrep[rownum,'right daughter'],height.increment)
    rval <- list(left,right)

    attr(rval,"members") <- attr(left,"members") + attr(right,"members")
    attr(rval,"height") <- max(attr(left,"height"),attr(right,"height")) + height.increment
    attr(rval,"leaf") <- FALSE
    attr(rval,"edgetext") <- dfrep[rownum,'split var']
    #To add Split Point in Dendrogram
    #attr(rval,"edgetext") <- paste(dfrep[rownum,'split var'],"\n<",round(dfrep[rownum,'split point'], digits = 2),"=>", sep = " ")
  }

  class(rval) <- "dendrogram"

  return(rval)
}

# Source own functions
# Define direwctory where functions are for each operating system

if (Sys.info()["sysname"] == 'Windows') {
  fun.dir <- 'D:/matchups/r-projects/Matchup_R_Scripts/Functions/'
} else if (Sys.info()["sysname"] == 'Linux') {
  fun.dir <- '/home/ckk/Projects/Matchup_R_Scripts/Functions/'
}

fun.file <- paste0(fun.dir, 'common_functions.R')

source(file = fun.file,
  local = FALSE, echo = FALSE, verbose = FALSE)
rm(fun.dir, fun.file)


# Load data
# For 787
linux_dir <- '~/Projects/Matchup_R_Scripts/Results/objects/'
# For Laptop
#linux_dir <- '~/Projects/Matchup_R_Scripts/'
linux_file <- 'MODIS_Aqua_GSFC_ALL_Class_6.4.1_ao_2017_04_12_with_ancillary.Rdata'
AQUA_file <- paste0(linux_dir, linux_file)

if (!file.exists(AQUA_file)) {
    stop('Input file does not exist')
  } else {
    load(AQUA_file, verbose = TRUE)
    AQUA <- MODIS_Aqua_GSFC_ALL_Class_6.4.1_ao_2017_04_12
  }

rm(linux_dir, linux_file, AQUA_file)
rm(MODIS_Aqua_GSFC_ALL_Class_6.4.1_ao_2017_04_12)

set.seed(42)
```

The matchup file read in at a previous step is not filtered and contains too many variables. For the purpose of 
this analysis, we use only nighttime matchups with a quality of 0, 1, or 2 and only keep variables from the infrared
channels we use.

```{r prep_data, echo=TRUE}
# Turn POSIXct objects to characters bc dplyr doesn't support POSIXct
AQUA$sat.timedate <- as.character(AQUA$sat.timedate)
AQUA$buoy.timedate <- as.character(AQUA$buoy.timedate)

# Apply basic filtering
AQUA <- dplyr::tbl_df(AQUA) %>%
  dplyr::filter(solz >= 90) %>%
  dplyr::filter(qsst == 0 | qsst == 1 | qsst == 2)

# Now grab only variables that may be used to determine retrieval accuracy and make some new variables (i.e. x1, x2, x3)
# Create df of features
orig <- dplyr::tbl_df(AQUA) %>%
  dplyr::mutate(T_11 = cen.11000,
    T_12 = cen.12000,
    band.diff = T_11 - T_12,
    ref_SST = cen.ref.type.1.SST,
    x2 = band.diff * ref_SST,
    x3 = ((secant.deg(satz) - 1) * band.diff),
    lat = buoy.lat,
    lon = buoy.lon,
    sd11 = sd.11000,
    sd12 = sd.12000,
    range11 = max.11000 - min.11000,
    range12 = max.12000 - min.12000,
    diff.med.min11 = med.11000 - min.11000,
    diff.med.min12 = med.12000 - min.12000,
    qsst = qsst,
    buoy.sst = buoy.sst,
    cell5deg = cell5deg,
    buoy.timedate = buoy.timedate,
    SST.resid.SMB = cen.sst - (buoy.sst - 0.17)) %>% # SMB = sat minus buoy - also debias buoy.sst by turning buoy.sst into a skin measurement
  dplyr::select(T_11, T_12, band.diff, ref_SST, x2, x3, satz, lon, lat, sd11, sd12, range11, range12, diff.med.min11, diff.med.min12,
    qsst, buoy.sst, cell5deg, buoy.timedate, SST.resid.SMB)

```


# Increased Spread of Residuals at Higher Buoy SSTs
We are interested in exploring the combinations of geophysical conditions that impact residual magnitude.
For starters, we plot the buoy SST against the SST residual and see that the spread of the residual greatly
increases at higher SST values.

```{r residuals_vs_buoy_SST, echo = TRUE}
ccc = RColorBrewer::brewer.pal(n = 8, 'YlOrRd')
ccc <- ccc[3:8]

ggplot2::ggplot(data = orig, mapping = aes(x = buoy.sst, y = SST.resid.SMB,
  fill = cut(..count.., c(0, 10, 50, 100, 250, 500, 1000, 2500, 5000, Inf)))) +
  geom_bin2d(bins = 75) +
  #scale_fill_gradientn(colours = ccc) +
  scale_fill_hue('count') +
  geom_abline(slope = 0, intercept = 0, col = 'black')

```

# Relationship Between band_diff and temp_deficit

## SST as a Separator
```{r band_diff_temp_deficit_context_SST, echo=TRUE}

###### NOTE
## ASK GUILLERMO HOW TO MAKE KEY FOR THE DIFFERENT LINES


zz1 <- dplyr::tbl_df(orig) %>%
  dplyr::filter(qsst == 0)

ccc = RColorBrewer::brewer.pal(n = 8, 'YlOrRd')
ccc <- ccc[3:8]

temp_context <- cut(zz1$buoy.sst, c(-Inf, 10, 25, Inf))
temp_deficit <- zz1$buoy.sst - zz1$T_11

uuu <- data.frame(temp_context = as.factor(temp_context), temp_deficit = temp_deficit, band.diff = zz1$band.diff)

# Take a sample of uuu bc loess requires lots of memory - for testing purposes
set.seed(42)
index <- sample(1:nrow(uuu), 50000, replace = FALSE)
uuu <- uuu[index, ]

ggplot2::ggplot(data = uuu, mapping = aes(x = band.diff, y = temp_deficit)) +
  geom_bin2d(bins = 20) +
  scale_fill_gradientn(colours = ccc) +
  geom_smooth(method = lm, se = FALSE, color = 'green', alpha = 0.1) +
  geom_smooth(method = loess, se = FALSE, span = 0.4) +
  facet_wrap(~temp_context, nrow = 3, ncol = 1) +
  labs(x = 'T_11 - T_12', y = 'Temperature Deficit')
  #ggsave('band_diff_predictor_of_temp_deficit_different_in_situ_windows.pdf')

# Correlation of each subset
## SST < 10
cat('First pearson and then spearman non parametric correlation coefs for SST <= 10')
cor(uuu[uuu$temp_context == '(-Inf,10]', "band.diff"], uuu[uuu$temp_context == '(-Inf,10]', "temp_deficit"])
cor.test(uuu[uuu$temp_context == '(-Inf,10]', "band.diff"],
         uuu[uuu$temp_context == '(-Inf,10]', "temp_deficit"],
         method = 'spearman', continuity = FALSE, conf.level = 0.95)

## 10 < SST < 25
cat('First pearson and then spearman non parametric correlation coefs for 10 < SST <= 25')
cor(uuu[uuu$temp_context == '(10,25]', "band.diff"], uuu[uuu$temp_context == '(10,25]', "temp_deficit"])
cor.test(uuu[uuu$temp_context == '(10,25]', "band.diff"],
         uuu[uuu$temp_context == '(10,25]', "temp_deficit"],
         method = 'spearman', continuity = FALSE, conf.level = 0.95)

## 25 < SST
cat('First pearson and then spearman non parametric correlation coefs for SST > 25')
cor(uuu[uuu$temp_context == '(25, Inf]', "band.diff"], uuu[uuu$temp_context == '(25, Inf]', "temp_deficit"])
cor.test(uuu[uuu$temp_context == '(25, Inf]', "band.diff"],
         uuu[uuu$temp_context == '(25, Inf]', "temp_deficit"],
         method = 'spearman', continuity = FALSE, conf.level = 0.95)


# Correlation of whole dataset
cor(uuu$band.diff, uuu$temp_deficit)
cor.test(uuu$band.diff, uuu$temp_deficit, method = 'spearman', continuity = FALSE, conf.level = 0.95)

```

## Residual as a Separator
```{r band_diff_temp_deficit_context_residual_qsst0_only, echo=TRUE}

zz1 <- dplyr::tbl_df(orig) %>%
  dplyr::filter(qsst == 0)

ccc = RColorBrewer::brewer.pal(n = 8, 'YlOrRd')
ccc <- ccc[3:8]

resid_context <- cut(zz1$SST.resid.SMB, c(-Inf, -.4, .4, Inf))
temp_deficit <- zz1$buoy.sst - zz1$T_11

uuu <- data.frame(resid_context = as.factor(resid_context), temp_deficit = temp_deficit, band.diff = zz1$band.diff)

uuu <- dplyr::tbl_df(uuu) %>%
  dplyr::filter(resid_context != '(0.4, Inf]')

uuu <- as.data.frame(uuu)

# Take a sample of uuu bc loess requires lots of memory - for testing purposes
set.seed(42)
index <- sample(1:nrow(uuu), 50000, replace = FALSE)
uuu <- uuu[index, ]


ggplot2::ggplot(data = uuu, mapping = aes(x = band.diff, y = temp_deficit)) +
  geom_density2d() +
  #geom_bin2d(bins = 20) +
  #scale_fill_gradientn(colours = ccc) +
  geom_smooth(method = lm, se = FALSE, color = 'green', alpha = 0.1) +
  geom_smooth(method = loess, se = FALSE, span = 0.4) +
  facet_wrap(~resid_context, nrow = 3, ncol = 1) +
  labs(x = 'T_11 - T_12', y = 'Temperature Deficit')
  #ggsave('band_diff_predictor_of_temp_deficit_different_in_situ_windows.pdf')

# Correlation of each subset
## SST < 10
cat('First pearson and then spearman non parametric correlation coefs for resid <= -0.4')
cor(uuu[uuu$resid_context == '(-Inf,-0.4]', "band.diff"], uuu[uuu$resid_context == '(-Inf,-0.4]', "temp_deficit"])
cor.test(uuu[uuu$resid_context == '(-Inf,-0.4]', "band.diff"],
         uuu[uuu$resid_context == '(-Inf,-0.4]', "temp_deficit"],
         method = 'spearman', continuity = FALSE, conf.level = 0.95)

## 10 < SST < 25
cat('First pearson and then spearman non parametric correlation coefs for -0.4 < resid <= 0.4')
cor(uuu[uuu$resid_context == '(-0.4,0.4]', "band.diff"], uuu[uuu$resid_context == '(-0.4,0.4]', "temp_deficit"])
cor.test(uuu[uuu$resid_context == '(-0.4,0.4]', "band.diff"],
         uuu[uuu$resid_context == '(-0.4,0.4]', "temp_deficit"],
         method = 'spearman', continuity = FALSE, conf.level = 0.95)

## 25 < SST
#cat('First pearson and then spearman non parametric correlation coefs for resid > 0.4')
#cor(uuu[uuu$resid_context == '(0.4, Inf]', "band.diff"], uuu[uuu$resid_context == '(0.4, Inf]', "temp_deficit"])
#cor.test(uuu[uuu$resid_context == '(0.4, Inf]', "band.diff"],
#         uuu[uuu$resid_context == '(0.4, Inf]', "temp_deficit"],
#         method = 'spearman', continuity = FALSE, conf.level = 0.95)


# Correlation of whole dataset
cor(uuu$band.diff, uuu$temp_deficit)
cor.test(uuu$band.diff, uuu$temp_deficit, method = 'spearman', continuity = FALSE, conf.level = 0.95)
```






Same plot as above but qsst == 0, 1, or 2.
```{r band_diff_temp_deficit_context_residual_all_qualities, echo=TRUE}

zz1 <- dplyr::tbl_df(orig) %>%
  dplyr::filter(qsst != 3) %>%
  dplyr::filter(qsst != 4)

ccc = RColorBrewer::brewer.pal(n = 8, 'YlOrRd')
ccc <- ccc[3:8]

resid_context <- cut(zz1$SST.resid.SMB, c(-Inf, -.4, .4, Inf))
temp_deficit <- zz1$buoy.sst - zz1$T_11

uuu <- data.frame(resid_context = as.factor(resid_context), temp_deficit = temp_deficit, band.diff = zz1$band.diff)

# Take a sample of uuu bc loess requires lots of memory - for testing purposes
#set.seed(42)
#index <- sample(1:nrow(uuu), 50000, replace = FALSE)
#uuu <- uuu[index, ]


ggplot2::ggplot(data = uuu, mapping = aes(x = band.diff, y = temp_deficit)) +
  geom_bin2d(bins = 20) +
  scale_fill_gradientn(colours = ccc) +
  #geom_smooth(method = lm, se = FALSE, color = 'green', alpha = 0.1) +
  #geom_smooth(method = loess, se = FALSE, span = 0.4) +
  facet_wrap(~resid_context, nrow = 3, ncol = 1) +
  labs(x = 'T_11 - T_12', y = 'Temperature Deficit')
  #ggsave('band_diff_predictor_of_temp_deficit_different_in_situ_windows.pdf')

# Correlation of each subset
## SST < 10
cat('First pearson and then spearman non parametric correlation coefs for resid <= -0.4')
cor(uuu[uuu$resid_context == '(-Inf,-0.4]', "band.diff"], uuu[uuu$resid_context == '(-Inf,-0.4]', "temp_deficit"])
cor.test(uuu[uuu$resid_context == '(-Inf,-0.4]', "band.diff"],
         uuu[uuu$resid_context == '(-Inf,-0.4]', "temp_deficit"],
         method = 'spearman', continuity = FALSE, conf.level = 0.95)

## 10 < SST < 25
cat('First pearson and then spearman non parametric correlation coefs for -0.4 < resid <= 0.4')
cor(uuu[uuu$resid_context == '(-0.4,0.4]', "band.diff"], uuu[uuu$resid_context == '(-0.4,0.4]', "temp_deficit"])
cor.test(uuu[uuu$resid_context == '(-0.4,0.4]', "band.diff"],
         uuu[uuu$resid_context == '(-0.4,0.4]', "temp_deficit"],
         method = 'spearman', continuity = FALSE, conf.level = 0.95)

## 25 < SST
cat('First pearson and then spearman non parametric correlation coefs for resid > 0.4')
cor(uuu[uuu$resid_context == '(0.4, Inf]', "band.diff"], uuu[uuu$resid_context == '(0.4, Inf]', "temp_deficit"])
cor.test(uuu[uuu$resid_context == '(0.4, Inf]', "band.diff"],
         uuu[uuu$resid_context == '(0.4, Inf]', "temp_deficit"],
         method = 'spearman', continuity = FALSE, conf.level = 0.95)


# Correlation of whole dataset
cor(uuu$band.diff, uuu$temp_deficit)
cor.test(uuu$band.diff, uuu$temp_deficit, method = 'spearman', continuity = FALSE, conf.level = 0.95)

```

# Applying the NLSST Correction

NLSST accounts for the loss in linear correlation between the band difference and temperature deficit by
multiplying the band difference by the reference SST. We now explore that correlation.

## SST as a Separator
```{r band_diff_product_temp_deficit_context_sst, echo=TRUE}


zz1 <- dplyr::tbl_df(orig) %>%
  dplyr::filter(qsst == 0)

ccc = RColorBrewer::brewer.pal(n = 8, 'YlOrRd')
ccc <- ccc[3:8]

temp_context <- cut(zz1$buoy.sst, c(-Inf, 10, 25, Inf))
temp_deficit <- zz1$buoy.sst - zz1$T_11

uuu <- data.frame(temp_context = as.factor(temp_context), temp_deficit = temp_deficit, product = (zz1$band.diff * zz1$ref_SST))

# Take a sample of uuu bc loess requires lots of memory - for testing purposes
set.seed(42)
index <- sample(1:nrow(uuu), 50000, replace = FALSE)
uuu <- uuu[index, ]

ggplot2::ggplot(data = uuu, mapping = aes(x = product, y = temp_deficit)) +
  geom_bin2d(bins = 20) +
  scale_fill_gradientn(colours = ccc) +
  geom_smooth(method = lm, se = FALSE, color = 'green', alpha = 0.1) +
  geom_smooth(method = loess, se = FALSE, span = 0.4) +
  facet_wrap(~temp_context, nrow = 3, ncol = 1) +
  labs(x = '(T_11 - T_12) * Reference SST', y = 'Temperature Deficit')
  #ggsave('band_diff_predictor_of_temp_deficit_different_in_situ_windows.pdf')

# Correlation of each subset
## SST < 10
cat('First pearson and then spearman non parametric correlation coefs for SST <= 10')
cor(uuu[uuu$temp_context == '(-Inf,10]', "product"], uuu[uuu$temp_context == '(-Inf,10]', "temp_deficit"])
cor.test(uuu[uuu$temp_context == '(-Inf,10]', "product"],
         uuu[uuu$temp_context == '(-Inf,10]', "temp_deficit"],
         method = 'spearman', continuity = FALSE, conf.level = 0.95)

## 10 < SST < 25
cat('First pearson and then spearman non parametric correlation coefs for 10 < SST <= 25')
cor(uuu[uuu$temp_context == '(10,25]', "product"], uuu[uuu$temp_context == '(10,25]', "temp_deficit"])
cor.test(uuu[uuu$temp_context == '(10,25]', "product"],
         uuu[uuu$temp_context == '(10,25]', "temp_deficit"],
         method = 'spearman', continuity = FALSE, conf.level = 0.95)

## 25 < SST
cat('First pearson and then spearman non parametric correlation coefs for SST > 25')
cor(uuu[uuu$temp_context == '(25, Inf]', "product"], uuu[uuu$temp_context == '(25, Inf]', "temp_deficit"])
cor.test(uuu[uuu$temp_context == '(25, Inf]', "product"],
         uuu[uuu$temp_context == '(25, Inf]', "temp_deficit"],
         method = 'spearman', continuity = FALSE, conf.level = 0.95)


# Correlation of whole dataset
cor(uuu$product, uuu$temp_deficit)
cor.test(uuu$product, uuu$temp_deficit, method = 'spearman', continuity = FALSE, conf.level = 0.95)

```

## Residual as a Separator
```{r band_diff_product_temp_deficit_context_residual_qsst0_only, echo=TRUE}

zz1 <- dplyr::tbl_df(orig) %>%
  dplyr::filter(qsst == 0)

ccc = RColorBrewer::brewer.pal(n = 8, 'YlOrRd')
ccc <- ccc[3:8]

resid_context <- cut(zz1$SST.resid.SMB, c(-Inf, -.4, .4, Inf))
temp_deficit <- zz1$buoy.sst - zz1$T_11

uuu <- data.frame(resid_context = as.factor(resid_context), temp_deficit = temp_deficit, product = (zz1$band.diff * zz1$ref_SST))

# Take a sample of uuu bc loess requires lots of memory - for testing purposes
set.seed(42)
index <- sample(1:nrow(uuu), 50000, replace = FALSE)
uuu <- uuu[index, ]


ggplot2::ggplot(data = uuu, mapping = aes(x = product, y = temp_deficit)) +
  geom_bin2d(bins = 20) +
  scale_fill_gradientn(colours = ccc) +
  geom_smooth(method = lm, se = FALSE, color = 'green', alpha = 0.1) +
  geom_smooth(method = loess, se = FALSE, span = 0.4) +
  facet_wrap(~resid_context, nrow = 3, ncol = 1) +
  labs(x = '(T_11 - T_12) * Reference SST', y = 'Temperature Deficit')
  #ggsave('band_diff_predictor_of_temp_deficit_different_in_situ_windows.pdf')

# Correlation of each subset
## SST < 10
cat('First pearson and then spearman non parametric correlation coefs for resid <= -0.4')
cor(uuu[uuu$resid_context == '(-Inf,-0.4]', "product"], uuu[uuu$resid_context == '(-Inf,-0.4]', "temp_deficit"])
cor.test(uuu[uuu$resid_context == '(-Inf,-0.4]', "product"],
         uuu[uuu$resid_context == '(-Inf,-0.4]', "temp_deficit"],
         method = 'spearman', continuity = FALSE, conf.level = 0.95)

## 10 < SST < 25
cat('First pearson and then spearman non parametric correlation coefs for -0.4 < resid <= 0.4')
cor(uuu[uuu$resid_context == '(-0.4,0.4]', "product"], uuu[uuu$resid_context == '(-0.4,0.4]', "temp_deficit"])
cor.test(uuu[uuu$resid_context == '(-0.4,0.4]', "product"],
         uuu[uuu$resid_context == '(-0.4,0.4]', "temp_deficit"],
         method = 'spearman', continuity = FALSE, conf.level = 0.95)

## 25 < SST
cat('First pearson and then spearman non parametric correlation coefs for resid > 0.4')
cor(uuu[uuu$resid_context == '(0.4, Inf]', "product"], uuu[uuu$resid_context == '(0.4, Inf]', "temp_deficit"])
cor.test(uuu[uuu$resid_context == '(0.4, Inf]', "product"],
         uuu[uuu$resid_context == '(0.4, Inf]', "temp_deficit"],
         method = 'spearman', continuity = FALSE, conf.level = 0.95)


# Correlation of whole dataset
cor(uuu$product, uuu$temp_deficit)
cor.test(uuu$product, uuu$temp_deficit, method = 'spearman', continuity = FALSE, conf.level = 0.95)

```

Same plot as above but qsst == 0, 1, or 2.
```{r band_diff_product_temp_deficit_context_residual_all_qualities, echo=TRUE}

zz1 <- dplyr::tbl_df(orig) %>%
  dplyr::filter(qsst != 3) %>%
  dplyr::filter(qsst != 4)

ccc = RColorBrewer::brewer.pal(n = 8, 'YlOrRd')
ccc <- ccc[3:8]

resid_context <- cut(zz1$SST.resid.SMB, c(-Inf, -.4, .4, Inf))
temp_deficit <- zz1$buoy.sst - zz1$T_11

uuu <- data.frame(resid_context = as.factor(resid_context), temp_deficit = temp_deficit, product = (zz1$band.diff * zz1$ref_SST))

# Take a sample of uuu bc loess requires lots of memory - for testing purposes
set.seed(42)
index <- sample(1:nrow(uuu), 50000, replace = FALSE)
uuu <- uuu[index, ]


ggplot2::ggplot(data = uuu, mapping = aes(x = product, y = temp_deficit)) +
  geom_bin2d(bins = 20) +
  scale_fill_gradientn(colours = ccc) +
  geom_smooth(method = lm, se = FALSE, color = 'green', alpha = 0.1) +
  geom_smooth(method = loess, se = FALSE, span = 0.4) +
  facet_wrap(~resid_context, nrow = 3, ncol = 1) +
  labs(x = 'T_11 - T_12', y = 'Temperature Deficit')
  #ggsave('band_diff_predictor_of_temp_deficit_different_in_situ_windows.pdf')

# Correlation of each subset
## SST < 10
cat('First pearson and then spearman non parametric correlation coefs for resid <= -0.4')
cor(uuu[uuu$resid_context == '(-Inf,-0.4]', "product"], uuu[uuu$resid_context == '(-Inf,-0.4]', "temp_deficit"])
cor.test(uuu[uuu$resid_context == '(-Inf,-0.4]', "product"],
         uuu[uuu$resid_context == '(-Inf,-0.4]', "temp_deficit"],
         method = 'spearman', continuity = FALSE, conf.level = 0.95)

## 10 < SST < 25
cat('First pearson and then spearman non parametric correlation coefs for -0.4 < resid <= 0.4')
cor(uuu[uuu$resid_context == '(-0.4,0.4]', "product"], uuu[uuu$resid_context == '(-0.4,0.4]', "temp_deficit"])
cor.test(uuu[uuu$resid_context == '(-0.4,0.4]', "product"],
         uuu[uuu$resid_context == '(-0.4,0.4]', "temp_deficit"],
         method = 'spearman', continuity = FALSE, conf.level = 0.95)

## 25 < SST
cat('First pearson and then spearman non parametric correlation coefs for resid > 0.4')
cor(uuu[uuu$resid_context == '(0.4, Inf]', "product"], uuu[uuu$resid_context == '(0.4, Inf]', "temp_deficit"])
cor.test(uuu[uuu$resid_context == '(0.4, Inf]', "product"],
         uuu[uuu$resid_context == '(0.4, Inf]', "temp_deficit"],
         method = 'spearman', continuity = FALSE, conf.level = 0.95)


# Correlation of whole dataset
cor(uuu$product, uuu$temp_deficit)
cor.test(uuu$product, uuu$temp_deficit, method = 'spearman', continuity = FALSE, conf.level = 0.95)

```

# No Separators

## Temp Deficit with Varying Buoy SST
```{r temp_deficit_buoy_SST_qsst0, echo=TRUE}

zz1 <- dplyr::tbl_df(orig) %>%
  dplyr::filter(qsst == 0)

ccc = RColorBrewer::brewer.pal(n = 8, 'YlOrRd')
ccc <- ccc[3:8]

temp_deficit <- zz1$buoy.sst - zz1$T_11

# Take a sample of zz1 bc loess requires lots of memory - for testing purposes
set.seed(42)
index <- sample(1:nrow(zz1), 50000, replace = FALSE)
zz1 <- zz1[index, ]

ggplot2::ggplot(mapping = aes(x = zz1$buoy.sst, y = temp_deficit[index])) +
  geom_bin2d(bins = 20) +
  scale_fill_gradientn(colours = ccc) +
  geom_smooth(method = lm, se = FALSE, color = 'green', alpha = 0.1) +
  geom_smooth(method = loess, se = FALSE, span = 0.4) +
  labs(x = 'Buoy SST', y = 'Temperature Deficit')
```


```{r temp_deficit_buoy_SST_qsst_all, echo=TRUE}

zz1 <- dplyr::tbl_df(orig) %>%
  dplyr::filter(qsst != 3) %>%
  dplyr::filter(qsst != 4)

ccc = RColorBrewer::brewer.pal(n = 8, 'YlOrRd')
ccc <- ccc[3:8]

temp_deficit <- zz1$buoy.sst - zz1$T_11

# Take a sample of zz1 bc loess requires lots of memory - for testing purposes
set.seed(42)
index <- sample(1:nrow(zz1), 50000, replace = FALSE)
zz1 <- zz1[index, ]

ggplot2::ggplot(mapping = aes(x = zz1$buoy.sst, y = temp_deficit[index])) +
  geom_bin2d(bins = 20) +
  scale_fill_gradientn(colours = ccc) +
  geom_smooth(method = lm, se = FALSE, color = 'green', alpha = 0.1) +
  geom_smooth(method = loess, se = FALSE, span = 0.4) +
  labs(x = 'Buoy SST', y = 'Temperature Deficit')
```

## Band Diff with Varying Buoy SST
```{r band_diff_buoy_SST_qsst0, echo=TRUE}

zz1 <- dplyr::tbl_df(orig) %>%
  dplyr::filter(qsst == 0)

ccc = RColorBrewer::brewer.pal(n = 8, 'YlOrRd')
ccc <- ccc[3:8]

temp_deficit <- zz1$buoy.sst - zz1$T_11

# Take a sample of zz1 bc loess requires lots of memory - for testing purposes
set.seed(42)
index <- sample(1:nrow(zz1), 50000, replace = FALSE)
zz1 <- zz1[index, ]

ggplot2::ggplot(mapping = aes(x = zz1$buoy.sst, y = zz1$band.diff)) +
  geom_bin2d(bins = 20) +
  scale_fill_gradientn(colours = ccc) +
  geom_smooth(method = lm, se = FALSE, color = 'green', alpha = 0.1) +
  geom_smooth(method = loess, se = FALSE, span = 0.4) +
  labs(x = 'Buoy SST', y = 'T_11 - T_12')
```


```{r band_diff_buoy_SST_qsst_all, echo=TRUE}

zz1 <- dplyr::tbl_df(orig) %>%
  dplyr::filter(qsst != 3) %>%
  dplyr::filter(qsst != 4)

ccc = RColorBrewer::brewer.pal(n = 8, 'YlOrRd')
ccc <- ccc[3:8]

temp_deficit <- zz1$buoy.sst - zz1$T_11

# Take a sample of zz1 bc loess requires lots of memory - for testing purposes
set.seed(42)
index <- sample(1:nrow(zz1), 50000, replace = FALSE)
zz1 <- zz1[index, ]

ggplot2::ggplot(mapping = aes(x = zz1$buoy.sst, y = zz1$band.diff)) +
  geom_bin2d(bins = 20) +
  scale_fill_gradientn(colours = ccc) +
  geom_smooth(method = lm, se = FALSE, color = 'green', alpha = 0.1) +
  geom_smooth(method = loess, se = FALSE, span = 0.4) +
  labs(x = 'Buoy SST', y = 'T_11 - T_12')
```

## Temp Deficit with Band Diff

```{r band_diff_temp_deficit_qsst0, echo=TRUE}

zz1 <- dplyr::tbl_df(orig) %>%
  dplyr::filter(qsst == 0)

ccc = RColorBrewer::brewer.pal(n = 8, 'YlOrRd')
ccc <- ccc[3:8]

temp_deficit <- zz1$buoy.sst - zz1$T_11

# Take a sample of zz1 bc loess requires lots of memory - for testing purposes
set.seed(42)
index <- sample(1:nrow(zz1), 50000, replace = FALSE)
zz1 <- zz1[index, ]

ggplot2::ggplot(mapping = aes(x = zz1$band.diff, y = temp_deficit[index])) +
  geom_bin2d(bins = 20) +
  scale_fill_gradientn(colours = ccc) +
  geom_smooth(method = lm, se = FALSE, color = 'green', alpha = 0.1) +
  geom_smooth(method = loess, se = FALSE, span = 0.4) +
  labs(x = 'T_11 - T_12', y = 'Temperature Deficit')
```


```{r band_diff_temp_deficit_qsst_all, echo=TRUE}

zz1 <- dplyr::tbl_df(orig) %>%
  dplyr::filter(qsst != 3) %>%
  dplyr::filter(qsst != 4)

ccc = RColorBrewer::brewer.pal(n = 8, 'YlOrRd')
ccc <- ccc[3:8]

temp_deficit <- zz1$buoy.sst - zz1$T_11

# Take a sample of zz1 bc loess requires lots of memory - for testing purposes
set.seed(42)
index <- sample(1:nrow(zz1), 50000, replace = FALSE)
zz1 <- zz1[index, ]

ggplot2::ggplot(mapping = aes(x = zz1$band.diff, y = temp_deficit[index])) +
  geom_bin2d(bins = 20) +
  scale_fill_gradientn(colours = ccc) +
  geom_smooth(method = lm, se = FALSE, color = 'green', alpha = 0.1) +
  geom_smooth(method = loess, se = FALSE, span = 0.4) +
  labs(x = 'T_11 - T_12', y = 'Temperature Deficit')
```


## Temp Deficit with Product


```{r band_diff_temp_deficit_qsst0, echo=TRUE}

zz1 <- dplyr::tbl_df(orig) %>%
  dplyr::filter(qsst == 0)

ccc = RColorBrewer::brewer.pal(n = 8, 'YlOrRd')
ccc <- ccc[3:8]

temp_deficit <- zz1$buoy.sst - zz1$T_11

# Take a sample of zz1 bc loess requires lots of memory - for testing purposes
set.seed(42)
index <- sample(1:nrow(zz1), 50000, replace = FALSE)
zz1 <- zz1[index, ]

ggplot2::ggplot(mapping = aes(x = (zz1$band.diff * zz1$ref_SST), y = temp_deficit[index])) +
  geom_bin2d(bins = 20) +
  scale_fill_gradientn(colours = ccc) +
  geom_smooth(method = lm, se = FALSE, color = 'green', alpha = 0.1) +
  geom_smooth(method = loess, se = FALSE, span = 0.4) +
  labs(x = '(T_11 - T_12) * Reference SST', y = 'Temperature Deficit')
```


```{r band_diff_temp_deficit_qsst_all, echo=TRUE}

zz1 <- dplyr::tbl_df(orig) %>%
  dplyr::filter(qsst != 3) %>%
  dplyr::filter(qsst != 4)

ccc = RColorBrewer::brewer.pal(n = 8, 'YlOrRd')
ccc <- ccc[3:8]

temp_deficit <- zz1$buoy.sst - zz1$T_11

# Take a sample of zz1 bc loess requires lots of memory - for testing purposes
set.seed(42)
index <- sample(1:nrow(zz1), 50000, replace = FALSE)
zz1 <- zz1[index, ]

ggplot2::ggplot(mapping = aes(x = (zz1$band.diff * zz1$ref_SST), y = temp_deficit[index])) +
  geom_bin2d(bins = 20) +
  scale_fill_gradientn(colours = ccc) +
  geom_smooth(method = lm, se = FALSE, color = 'green', alpha = 0.1) +
  geom_smooth(method = loess, se = FALSE, span = 0.4) +
  labs(x = '(T_11 - T_12) * Reference SST', y = 'Temperature Deficit')
```

## Temp Deficit with Product and Color as Residual

```{r band_diff_temp_deficit_qsst0, echo=TRUE}

zz1 <- dplyr::tbl_df(orig) %>%
  dplyr::filter(qsst == 0)

ccc = RColorBrewer::brewer.pal(n = 8, 'YlOrRd')
ccc <- ccc[3:8]

temp_deficit <- zz1$buoy.sst - zz1$T_11

# Take a sample of zz1 bc loess requires lots of memory - for testing purposes
set.seed(42)
index <- sample(1:nrow(zz1), 50000, replace = FALSE)
zz1 <- zz1[index, ]

ggplot2::ggplot(mapping = aes(x = (zz1$band.diff * zz1$ref_SST), y = temp_deficit[index], color = zz1$SST.resid.SMB)) +
  geom_point(alpha = 0.4) +
  #geom_jitter() + 
  #scale_color_manual(values = wesanderson::wes_palette(n = 8, name = 'YlOrRd')) +
  geom_smooth(method = lm, se = FALSE, color = 'green', alpha = 0.1) +
  geom_smooth(method = loess, se = FALSE, span = 0.4) +
  labs(x = '(T_11 - T_12) * Reference SST', y = 'Temperature Deficit')
```


```{r band_diff_temp_deficit_qsst_all, echo=TRUE}

zz1 <- dplyr::tbl_df(orig) %>%
  dplyr::filter(qsst != 3) %>%
  dplyr::filter(qsst != 4)

ccc = RColorBrewer::brewer.pal(n = 8, 'YlOrRd')
ccc <- ccc[3:8]

temp_deficit <- zz1$buoy.sst - zz1$T_11

# Take a sample of zz1 bc loess requires lots of memory - for testing purposes
set.seed(42)
index <- sample(1:nrow(zz1), 50000, replace = FALSE)
zz1 <- zz1[index, ]

ggplot2::ggplot(mapping = aes(x = (zz1$band.diff * zz1$ref_SST), y = temp_deficit[index], color = zz1$SST.resid.SMB)) +
  geom_point(alpha = 0.4) +
  #geom_jitter() +
  #scale_fill_gradientn(colours = ccc) +
  geom_smooth(method = lm, se = FALSE, color = 'green', alpha = 0.1) +
  geom_smooth(method = loess, se = FALSE, span = 0.4) +
  labs(x = '(T_11 - T_12) * Reference SST', y = 'Temperature Deficit')
```