---
title: "Exploration of a MODIS SSES Dataset - Temperature Deficit"
author: "Chirag Kumar and Guillermo Podesta"
output:
  html_notebook:
    toc: True
    theme: united
---


# Prep Workspace and Data
We use MODIS matchups from 2002 - 2016 that have been read in through a previous script. Here we simply load
the matchups, any necessary packages, and define necessary functions.

```{r prep_workspace, include=FALSE}
# Import necessary packages
require(ggplot2)
require(dplyr)
require(RColorBrewer)
require(circular)
require(ggmap)
require(raster)
require(sfsmisc)
require(rasterVis)
require(rgdal)
require(rpart)
require(rpart.plot)
require(caret)
require(evtree)
require(ctree)
require(partykit)
require(randomForest)
require(RWeka)
require(DMwR)

ggplot <- function(...) {ggplot2::ggplot(...) + theme_bw()}

# Define secant function
secant.deg <- function(x) {1 / (cos(circular::rad(x)))}

# Define function to turn output of gettree from random forest into a dendrogram
to.dendrogram <- function(dfrep,rownum=1,height.increment=0.1){

  if(dfrep[rownum,'status'] == -1){
    rval <- list()

    attr(rval,"members") <- 1
    attr(rval,"height") <- 0.0
    attr(rval,"label") <- dfrep[rownum,'prediction']
    attr(rval,"leaf") <- TRUE

  }else{##note the change "to.dendrogram" and not "to.dendogram"
    left <- to.dendrogram(dfrep,dfrep[rownum,'left daughter'],height.increment)
    right <- to.dendrogram(dfrep,dfrep[rownum,'right daughter'],height.increment)
    rval <- list(left,right)

    attr(rval,"members") <- attr(left,"members") + attr(right,"members")
    attr(rval,"height") <- max(attr(left,"height"),attr(right,"height")) + height.increment
    attr(rval,"leaf") <- FALSE
    attr(rval,"edgetext") <- dfrep[rownum,'split var']
    #To add Split Point in Dendrogram
    #attr(rval,"edgetext") <- paste(dfrep[rownum,'split var'],"\n<",round(dfrep[rownum,'split point'], digits = 2),"=>", sep = " ")
  }

  class(rval) <- "dendrogram"

  return(rval)
}

# Source own functions
# Define direwctory where functions are for each operating system

if (Sys.info()["sysname"] == 'Windows') {
  fun.dir <- 'D:/matchups/r-projects/Matchup_R_Scripts/Functions/'
} else if (Sys.info()["sysname"] == 'Linux') {
  fun.dir <- '/home/ckk/Projects/Matchup_R_Scripts/Functions/'
}

fun.file <- paste0(fun.dir, 'common_functions.R')

source(file = fun.file,
  local = FALSE, echo = FALSE, verbose = FALSE)
rm(fun.dir, fun.file)


# Load data
# For 787
linux_dir <- '~/Projects/Matchup_R_Scripts/Results/objects/'
# For Laptop
#linux_dir <- '~/Projects/Matchup_R_Scripts/'
linux_file <- 'MODIS_Aqua_GSFC_ALL_Class_6.4.1_ao_2017_04_12_with_ancillary.Rdata'
AQUA_file <- paste0(linux_dir, linux_file)

if (!file.exists(AQUA_file)) {
    stop('Input file does not exist')
  } else {
    load(AQUA_file, verbose = TRUE)
    AQUA <- MODIS_Aqua_GSFC_ALL_Class_6.4.1_ao_2017_04_12
  }

rm(linux_dir, linux_file, AQUA_file)
rm(MODIS_Aqua_GSFC_ALL_Class_6.4.1_ao_2017_04_12)

set.seed(42)
```

The matchup file read in at a previous step is not filtered and contains too many variables. For the purpose of 
this analysis, we use only nighttime matchups with a quality of 0, 1, or 2 and only keep variables from the infrared
channels we use.

```{r prep_data, echo=TRUE}
# Turn POSIXct objects to characters bc dplyr doesn't support POSIXct
AQUA$sat.timedate <- as.character(AQUA$sat.timedate)
AQUA$buoy.timedate <- as.character(AQUA$buoy.timedate)

# Apply basic filtering
AQUA <- dplyr::tbl_df(AQUA) %>%
  dplyr::filter(solz >= 90) %>%
  dplyr::filter(qsst == 0 | qsst == 1 | qsst == 2)

# Now grab only variables that may be used to determine retrieval accuracy and make some new variables (i.e. x1, x2, x3)
# Create df of features
orig <- dplyr::tbl_df(AQUA) %>%
  dplyr::mutate(T_11 = cen.11000,
    T_12 = cen.12000,
    band.diff = T_11 - T_12,
    ref_SST = cen.ref.type.1.SST,
    x2 = band.diff * ref_SST,
    x3 = ((secant.deg(satz) - 1) * band.diff),
    lat = buoy.lat,
    lon = buoy.lon,
    sd11 = sd.11000,
    sd12 = sd.12000,
    range11 = max.11000 - min.11000,
    range12 = max.12000 - min.12000,
    diff.med.min11 = med.11000 - min.11000,
    diff.med.min12 = med.12000 - min.12000,
    qsst = qsst,
    buoy.sst = buoy.sst,
    cell5deg = cell5deg,
    buoy.timedate = buoy.timedate,
    SST.resid.SMB = cen.sst - (buoy.sst - 0.17)) %>% # SMB = sat minus buoy - also debias buoy.sst by turning buoy.sst into a skin measurement
  dplyr::select(T_11, T_12, band.diff, ref_SST, x2, x3, satz, lon, lat, sd11, sd12, range11, range12, diff.med.min11, diff.med.min12,
    qsst, buoy.sst, cell5deg, buoy.timedate, SST.resid.SMB)

```


# Increased Spread of Residuals at Higher Buoy SSTs
We are interested in exploring the combinations of geophysical conditions that impact residual magnitude.
For starters, we plot the buoy SST against the SST residual.
```{r residuals_vs_buoy_SST, echo = TRUE}
ccc = RColorBrewer::brewer.pal(n = 8, 'YlOrRd')
ccc <- ccc[3:8]

ggplot2::ggplot(data = orig, mapping = aes(x = buoy.sst, y = SST.resid.SMB)) +
  geom_bin2d(bins = 100) +
  scale_fill_gradientn(colours = ccc)

residuals_buoy_sst_lm <- lm(SST.resid.SMB ~ buoy.sst, data = orig)

```

We see that for higher buoy SST values, the spread and magnitude of the residuals increases. At, say, 25 deg C this trend becomes even 
more evident where there is a clump of retrievals around residual = -2, which only happens when buoy SST exceeds 25.
We demonstrate that observation with the following statistics.

```{r table_residual_spread_for_buoy_sst_cutoffs, echo = TRUE}
ccc = RColorBrewer::brewer.pal(n = 8, 'YlOrRd')
ccc <- ccc[3:8]

ggplot2::ggplot(data = orig, mapping = aes(x = buoy.sst, y = SST.resid.SMB)) +
  geom_bin2d(bins = 100) +
  scale_fill_gradientn(colours = ccc)

residuals_buoy_sst_lm <- lm(SST.resid.SMB ~ buoy.sst, data = orig)

cutoff <- 23

summary(orig$SST.resid.SMB[orig$buoy.sst < cutoff])
```

Given that the residual's spread increases as the buoy SST increases, so should the temperature deficit.

```{r temp_deficit_vs_buoy_sst, echo = TRUE}

ccc = RColorBrewer::brewer.pal(n = 8, 'YlOrRd')
ccc <- ccc[3:8]

temp_deficit <- orig$buoy.sst - orig$x1

ggplot2::ggplot(data = orig, mapping = aes(x = buoy.sst, y = temp_deficit)) +
  geom_bin2d(bins = 100) +
  scale_fill_gradientn(colours = ccc)

temp_deficit_buoy_sst_lm <- lm(temp_deficit ~ buoy.sst, data = orig)

```

We have now observed, in multiple forms, that the residuals and temperature deficit increase in spread and magnitude as the
buoy SST increases.

In terms of identifying SST retrievals of varying residual magnitudes, figuring out what causes this spread and magnitude
increase would enable us to identify those retrievals.

# Motivation

We now explore possible geophysical conditions that could be impacting the increase in spread and magnitude of the residual
and temperature deficit we previously explored.

We use the band.diff (T_11 - T_12) as a proxy for atmospheric absorption. However, from the Clausius-Clapeyron
relationship, we know that at higher temperatures, the atmosphere can hold more water vapor. At higher SSTs, the
band.diff becomes less responsive and is no longer proportional to the temperature deficit.

```{r temp_deficit_vs_band_diff, echo = TRUE}

zz1 <- dplyr::tbl_df(orig) %>%
  dplyr::filter(qsst == 0)

ccc = RColorBrewer::brewer.pal(n = 8, 'YlOrRd')
ccc <- ccc[3:8]

temp_context <- cut(zz1$buoy.sst, c(-Inf, 10, 25, Inf))
temp_deficit <- zz1$buoy.sst - zz1$x1

uuu <- data.frame(temp_context = as.factor(temp_context), temp_deficit = temp_deficit, band.diff = zz1$band.diff)

# Take a sample of uuu bc loess requires lots of memory - for testing purposes
#set.seed(42)
#index <- sample(1:nrow(uuu), 1000, replace = FALSE)
#uuu <- uuu[index, ]

ggplot2::ggplot(data = uuu, aes(x = band.diff, y = temp_deficit)) +
  geom_bin2d(bins = 100) +
  scale_fill_gradientn(colours = ccc) + 
  geom_smooth(method = loess, se = FALSE, span = 0.5) +
  facet_wrap(~temp_context, nrow = 3, ncol = 1) +
  ggsave('band_diff_predictor_of_temp_deficit_different_in_situ_windows.pdf')

# Regression equations and statistics for the three facets
uuu$temp_context_numeric <- factor(as.character(uuu$temp_context), labels = c(0, 1, 2), levels = c('(-Inf,10]', '(10,25]', '(25, Inf]'))

temp_less_10_lm <- lm(formula = band.diff ~ temp_deficit, data = uuu[uuu$temp_context_numeric == 0, ])
rSD_temp_less_10 <- IQR(temp_less_10_lm$residuals) / 1.348
cor_temp_less_10 <- cor(uuu[uuu$temp_context_numeric == 0, 'temp_deficit'], uuu[uuu$temp_context_numeric == 0, 'band.diff'])

temp_between_10_25_lm <- lm(formula = band.diff ~ temp_deficit, data = uuu[uuu$temp_context_numeric == 1, ])
rSD_temp_between_10_25 <- IQR(temp_between_10_25_lm$residuals) / 1.348
cor_temp_between_10_25 <- cor(uuu[uuu$temp_context_numeric == 1, 'temp_deficit'], uuu[uuu$temp_context_numeric == 1, 'band.diff'])

temp_greater_25_lm <- lm(formula = band.diff ~ temp_deficit, data = uuu[uuu$temp_context_numeric == 2, ])
rSD_temp_greater_25 <- IQR(temp_greater_25_lm$residuals) / 1.348
cor_temp_greater_25 <- cor(uuu[uuu$temp_context_numeric == 2, 'temp_deficit'], uuu[uuu$temp_context_numeric == 2, 'band.diff'])

# Whole dataset for comparison
temp_all_lm <- lm(formula = band.diff ~ temp_deficit, data = uuu)
rSD_temp_all <- IQR(temp_all_lm$residuals) / 1.348
cor_temp_all <- cor(uuu$temp_deficit, uuu$band.diff)

```

To correct for this lose in proportionality at higher temperatures, the NLSST algorithm "scales" the band.diff by
the reference SST and the secant of the satellite zenith angle.

While these scaling measures help to create a term that is proportional to the temperature deficit, there is still
an issue...

```{r temp_deficit_vs_x2, echo = TRUE}

temp_deficit <- orig$buoy.sst - orig$x1
ggplot2::ggplot(data = orig, mapping = aes(x = x2, y = temp_deficit)) +
  geom_bin2d(bins = 100) +
  scale_fill_gradientn(colours = ccc)

```



```{r temp_deficit_vs_x3, echo = TRUE}

temp_deficit <- orig$buoy.sst - orig$x1
ggplot2::ggplot(data = orig, mapping = aes(x = x3, y = temp_deficit)) +
  geom_bin2d(bins = 100) +
  scale_fill_gradientn(colours = ccc)

```

# Empirical Densities

We now try to identify the combinations of conditions that lead to the retrievals with a larger residual magnitude
when SST is greater than 23 deg C.
We use empirical density estimates as our visualization tool.

```{r empirical_density_sd11, echo = TRUE}

qqq <- orig %>%
  dplyr::filter(buoy.sst >= 23)

ttt <- data.frame(condition = ifelse(qqq$sd11 > 2, ifelse(qqq$sd11 > 5, 2, 1), 0), resid = qqq$SST.resid.SMB)
qqq$condition <- factor(qqq$condition, levels = c(0, 1, 2), labels = c('sd11 less than 2', 'sd11 in-between 2 and 5', 'sd11 greater than 5'))

ggplot2::ggplot(data = ttt, aes(x = resid, fill = condition)) +
  geom_density(alpha = .4, color = NA) +
  geom_rug(aes(x = resid, y = 0), position = position_jitter(height = 0))
```


# Decision Trees





