---
title: "Exploration of a MODIS SSES Dataset - Empirical Densities"
author: "Chirag Kumar and Guillermo Podesta"
output:
  html_notebook:
    toc: True
    theme: united
---

In this notebook we explore characteristics of the SST residual, defined as satellite minus buoy SST, and what 
variables and conditions have significant impact on the residual's value and distribution. Our primary tool is
empirical densities where we explore the effects of conditions on residual density bias, spread, and shape.

We also explore the effect of SST on the relation between band_diff and temp_deficit a la Walton et al. 2016.
We expect that as band_diff increases, so will temp_deficit but this relationship will begin to plateau at high
values of band_diff. 

# Prep Workspace and Data
We use MODIS matchups from 2002 - 2016 that have been read in through a previous script. Here we simply load
the matchups, any necessary packages, and define necessary functions.

```{r prep_workspace, include=FALSE}
# Import necessary packages
require(ggplot2)
require(lubridate)
require(dplyr)
require(RColorBrewer)
require(circular)
require(ggmap)
require(raster)
require(sfsmisc)
require(rasterVis)
require(rgdal)

ggplot <- function(...) {ggplot2::ggplot(...) + theme_bw()}

# Define secant function
secant.deg <- function(x) {1 / (cos(circular::rad(x)))}

# Source own functions
# Define direwctory where functions are for each operating system

if (Sys.info()["sysname"] == 'Windows') {
  fun.dir <- 'D:/matchups/r-projects/Matchup_R_Scripts/Functions/'
} else if (Sys.info()["sysname"] == 'Linux') {
  fun.dir <- '/home/ckk/Projects/Matchup_R_Scripts/Functions/'
}

fun.file <- paste0(fun.dir, 'common_functions.R')

source(file = fun.file,
  local = FALSE, echo = FALSE, verbose = FALSE)
rm(fun.dir, fun.file)


# Load data
# For 787
linux_dir <- '~/Projects/Matchup_R_Scripts/Results/objects/'
# For Laptop
#linux_dir <- '~/Projects/Matchup_R_Scripts/'
linux_file <- 'MODIS_Aqua_GSFC_ALL_Class_6.4.1_ao_2017_04_12_with_ancillary.Rdata'
AQUA_file <- paste0(linux_dir, linux_file)

if (!file.exists(AQUA_file)) {
    stop('Input file does not exist')
  } else {
    load(AQUA_file, verbose = TRUE)
    AQUA <- MODIS_Aqua_GSFC_ALL_Class_6.4.1_ao_2017_04_12
  }

rm(linux_dir, linux_file, AQUA_file)
rm(MODIS_Aqua_GSFC_ALL_Class_6.4.1_ao_2017_04_12)
```

The matchup file read in at a previous step is not filtered and contains too many variables. For the purpose of 
this analysis, we use only nighttime matchups with a quality of 0, 1, or 2 and only keep variables from the infrared
channels we use.

```{r prep_data, echo=TRUE}
# Turn POSIXct objects to characters bc dplyr doesn't support POSIXct
AQUA$sat.timedate <- as.character(AQUA$sat.timedate)
AQUA$buoy.timedate <- as.character(AQUA$buoy.timedate)

# Apply basic filtering
AQUA <- dplyr::tbl_df(AQUA) %>%
  dplyr::filter(solz >= 90) %>%
  dplyr::filter(qsst == 0 | qsst == 1 | qsst == 2)

# Now grab only variables that may be used to determine retrieval accuracy and make some new variables (i.e. x1, x2, x3)
# Create df of features
orig <- dplyr::tbl_df(AQUA) %>%
  dplyr::mutate(x1 = cen.11000,
    x2 = cen.11000 - cen.12000,
    x3 = cen.ref.type.1.SST,
    satz = satz,
    x2 = x2 * x3,
    x3 = ((secant.deg(satz) - 1) * x2),
    lat = buoy.lat,
    lon = buoy.lon,
    sd11 = sd.11000,
    sd12 = sd.12000,
    range11 = max.11000 - min.11000,
    range12 = max.12000 - min.12000,
    diff.med.min11 = med.11000 - min.11000,
    diff.med.min12 = med.12000 - min.12000,
    band.diff = cen.11000 - cen.12000,
    qsst = qsst,
    buoy.sst = buoy.sst,
    cell5deg = cell5deg,
    buoy.timedate = buoy.timedate,
    SST.resid.SMB = cen.sst - buoy.sst) %>% # SMB = sat minus buoy - also don't debias SSTs i.e. turn buoy into skin
  dplyr::select(x1, x2, x3, satz, lon, lat, sd11, sd12, range11, range12, diff.med.min11, diff.med.min12, band.diff, qsst, buoy.sst, cell5deg,
                buoy.timedate, SST.resid.SMB)

```


# Spatial Distribution of Matchups

Before diving into our analysis of conditions that effect residual magnitude, we do some basic explorations of
our dataset. First we look at the spatial distribution of matchups


```{r spatial_distribution, echo=TRUE}
mp <- NULL
mapWorld <- borders("world", colour = "gray50", fill = "gray70") # create a layer of borders
mp <- ggplot() +  mapWorld +
  ggplot2::scale_x_continuous(breaks = seq(from = -180, to = 180, by = 60)) +
  ggplot2::scale_y_continuous(breaks = seq(from = -90, to = 90, by = 30)) +
  coord_fixed(ratio = 1)

# WARNING: Be careful as this worked with ggplot2 2.1.0 on Linux
# Try to work on different machines with varying versions of ggplot2
# Now layer the buoys on top with hexagonal binning
mp <- mp + 
  ggplot2::stat_binhex(data = orig,
    aes(x = lon, # Use hexagonal binning command and give x and y input
    y = lat,
    fill = cut(..count.., c(0, 1000, 2000, 5000, 10000, Inf))), # Divides matchups per bin into discrete chunks and colors likewise
    binwidth = c(10, 10)) +
  mapWorld + labs(x = NULL, y = NULL) +
  #scale_fill_hue('value') + # Standard colors with discrete chunking
  scale_fill_brewer(palette = 'YlOrRd') + # Change colors to Yellow, Orange, and Red - many diff 
  guides(fill = guide_legend(title = "N of matchups"))

#ggplot2::ggsave(filename = 'spatial_distribution.ps', device = 'ps',
#       width = 8, height = 6, units = 'in')

mp
```

```{r N_of_matchups_by_location_raster, echo=TRUE}
# --- These numbers may help perform statistics for the matchups
# --- (e.g., number of matchups per cell, etc.).

# --- Create a raster object with 5-degree pixels

crs.string <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"

grid5deg <- raster::raster(ncol = 72, nrow = 36,
  xmn = -180, xmx = 180,
  ymn = -90, ymx = 90,
  crs = crs.string)

# --- Write cell numbers as value for the grids.
# --- Cell 1 is the upper left corner and numbers go
# --- across the top line through its end, then
# --- start again in the second line, and so forth.

raster::values(grid5deg) <- 1:raster::ncell(grid5deg) # Cell numbers for 5 deg grid

raster::cellFromRowCol(object = grid5deg, rownr = 2, colnr = 72)

raster::xyFromCell(object =  grid5deg, cell = 2592, spatial=FALSE)

qqq <- dplyr::tbl_df(orig) %>%
  dplyr::group_by(cell5deg) %>%
  dplyr::summarise(N = n(), med = median(SST.resid.SMB), IQR = IQR(SST.resid.SMB))

qq1 <- rep(NA, (72*36))
qq1[qqq$cell5deg] <- qqq$N

raster::values(grid5deg) <- qq1

p <- rasterVis::levelplot(grid5deg, margin = FALSE,
  #at = c(-Inf, seq(from = -0.4, to = 0.4, by = 0.1), Inf), # For med
  at = c(1, 100, 250, 500, 1000, 2000, 5000, 10000, Inf), # For N
  col.regions = (RColorBrewer::brewer.pal(9, 'OrRd')))

world.map <- rgdal::readOGR("/home/ckk/Projects/Matchup_R_Scripts", layer = "ne_110m_land", verbose = TRUE)

p <- p + layer(sp.lines(world.map, lwd = .75, fill = 'grey70', col = 'yellow'))

p

```


# Temporal Distribution of Matchups

We now look at the temporal distribution of matchups

```{r heatmap_of_matchups, echo=TRUE}

yy <- lubridate::year(orig$buoy.timedate)     # Year
mm <- lubridate::month(orig$buoy.timedate)    # Month

tt1 <- cbind(yy, mm)

tt2 <- dplyr::tbl_df(tt1) %>%
  dplyr::group_by(yy, mm) %>%
  dplyr::summarise(n = n()) %>%
  dplyr::arrange(yy, mm)

yy.vals <- sort(unique(yy))
mm.vals <- sort(unique(mm))
tt3 <- expand.grid(mm = mm.vals, yy = yy.vals)

tt4 <- dplyr::left_join(tt3, tt2) %>%
  dplyr::select(yy, mm, n)

brks <- pretty(tt4$n, n = 5)
#brks <- floor(quantile(tt4$n, probs = seq(0, 1, 0.25), na.rm = TRUE))

tt4$n <- cut(tt4$n, breaks = brks)
tt4$mm <- ordered(tt4$mm, labels = month.abb)
tt4$yy <- ordered(tt4$yy)

hm <- ggplot2::ggplot(data = tt4, aes(mm, yy)) +
  ggplot2::geom_tile(aes(fill = n), colour = "white") +
  ggplot2::scale_fill_brewer(type = "seq", palette = 'YlOrRd', direction = 1) +
  ggplot2::theme_bw() +
  ggplot2::theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  ggplot2::ggsave("Temporal_heatmap.ps", device = 'ps',
       width = 8, height = 6, units = 'in')
hm
```


# Statistics of the SST Residual and Basic Exploration of the SST Residual

## Residuals by Quality
```{r resid_spread_quality, echo=TRUE}
# First look at statistics for the residuals in general
cat('\nSummary Statistics for All Residuals\n')
summary(orig$SST.resid.SMB)
cat('\n Standard Deviation for All Residuals: ')
sd(orig$SST.resid.SMB)

# Look at the resid spread among different quality levels
# qsst == 0
cat('\nSummary Statistics for Quality = 0\n')
summary(orig$SST.resid.SMB[orig$qsst == 0])
cat('\n Standard Deviation for Quality = 0: ')
sd(orig$SST.resid.SMB[orig$qsst == 0])

# qsst == 1
cat('\nSummary Statistics for Quality = 1\n')
summary(orig$SST.resid.SMB[orig$qsst == 1])
cat('\n Standard Deviation for Quality = 1: ')
sd(orig$SST.resid.SMB[orig$qsst == 1])

# qsst == 2
cat('\nSummary Statistics for Quality = 2\n')
summary(orig$SST.resid.SMB[orig$qsst == 2])
cat('\n Standard Deviation for Quality = 2: ')
sd(orig$SST.resid.SMB[orig$qsst == 2])
```

## Residuals as a Function of Buoy SST

```{r buoy_sst_vs_sst_resid, echo=TRUE}
ccc <- RColorBrewer::brewer.pal(n = 8, name = 'YlOrRd')
ccc <- ccc[3:8]

ggplot2::ggplot(data = orig, aes(x = buoy.sst, y = SST.resid.SMB + 0.17,
  fill = cut(..count.., c(0, 20, 100, 500, 1000, 1500, 2000, 4000, Inf)))) +
  geom_bin2d(bins = 100) +
  #scale_fill_gradientn(colours = ccc) +
  scale_fill_hue('count') +
  geom_abline(slope = 0, intercept = 0, col = 'black')

# Look at the statistics of the residuals
summary(orig$SST.resid.SMB)
IQR(orig$SST.resid.SMB)
```

## Residuals as a Function of Temperature Deficit

```{r sst_resid_vs_temp_deficit, echo=TRUE}
temp_deficit = orig$buoy.sst - orig$x1

ggplot2::ggplot(data = orig, aes(x = temp_deficit, y = SST.resid.SMB,
  fill = cut(..count.., c(0, 20, 100, 500, 1000, 1500, 2000, 4000, Inf)))) +
  geom_bin2d(bins = 100) + 
  scale_fill_hue('count') +
  geom_abline(slope = 0, intercept = 0, col = 'black')
```

## Residuals as a Function of Band Difference

```{r sst_resid_vs_band_diff, echo=TRUE}
ggplot2::ggplot(data = orig, aes(x = band.diff, y = SST.resid.SMB + 0.17,
  fill = cut(..count.., c(0, 20, 100, 500, 1000, 1500, 2000, 4000, Inf)))) +
  geom_bin2d(bins = 100) + 
  scale_fill_hue('count') +
  geom_abline(slope = 0, intercept = 0, col = 'black')
```


# Empirical Densities

```{r empirical_density_qsst, echo=TRUE}

ggplot2::ggplot(data = orig) + 
  geom_density(aes(x = SST.resid.SMB + 0.17, fill = as.factor(qsst)), alpha = 0.3, color = NA) +
  geom_rug(aes(x = SST.resid.SMB + 0.17, y = 0), position = position_jitter(height = 0))

```

```{r empirical_density_temp_deficit, echo=TRUE}

temp_deficit <- orig$buoy.sst - orig$x1
cutoff_val <- 15
qqq <- data.frame(condition = ifelse(temp_deficit > cutoff_val, paste('temp_deficit greater than', cutoff_val), paste('temp_deficit less than', cutoff_val)), resid = orig$SST.resid.SMB)

ggplot2::ggplot(data = qqq, aes(x = resid + 0.17, fill = condition)) +
  geom_density(alpha = .4, color = NA) +
  geom_rug(aes(x = SST.resid.SMB + 0.17, y = 0), position = position_jitter(height = 0))

```

```{r empirical_density_band_diff, echo=TRUE}

cutoff_val_down <- 0
cutoff_val_up <- 3
qqq <- data.frame(condition_band_diff = ifelse(orig$band.diff > cutoff_val_down, ifelse(orig$band.diff > cutoff_val_up, 2, 1), 0),
                  resid = orig$SST.resid.SMB + 0.17)

qqq$condition_band_diff <- factor(qqq$condition_band_diff, levels = c(0, 1, 2), labels = c('(-Inf, 0]', '(0, 2]', '(2, Inf)'))

ggplot2::ggplot(data = qqq, aes(x = resid + 0.17, fill = condition_band_diff)) +
  geom_density(alpha = .4, color = NA) +
  geom_rug(aes(x = SST.resid.SMB + 0.17, y = 0), position = position_jitter(height = 0)) +
  ggplot2::ggsave('band_diff_empirical_density.pdf')

```

```{r empirical_density_satz, echo=TRUE}

cutoff_val <- 55
qqq <- data.frame(condition = ifelse(orig$satz > cutoff_val, paste('satz greater than', cutoff_val), paste('satz less than', cutoff_val)), resid = orig$SST.resid.SMB)

ggplot2::ggplot(data = qqq, aes(x = resid + 0.17, fill = condition)) +
  geom_density(alpha = .4, color = NA) +
  geom_rug(aes(x = SST.resid.SMB + 0.17, y = 0), position = position_jitter(height = 0)) +
  ggplot2::ggsave('satz_empirical_density.pdf')

```

```{r empirical_density_band_diff_and_satz, echo=TRUE}

cutoff_val <- 0
qqq <- data.frame(condition_band_diff = ifelse(orig$band.diff > cutoff_val, ifelse(orig$band.diff > 3, 2, 1), 0),
                  condition_satz = cut(abs(orig$satz), c(0, 55, max(orig$satz) + .01), include.lowest = TRUE), # Somewhat arbitrary - To be revised
                  resid = orig$SST.resid.SMB + 0.17)

qqq <- qqq[!is.na(qqq$condition_satz), ]

qqq$condition_band_diff <- factor(qqq$condition_band_diff, levels = c(0, 1, 2), labels = c('(-Inf, 0]', '(0, 2]', '(2, Inf)'))

qqq$condition_total <- paste0('band_diff ', as.character(qqq$condition_band_diff), ' and satz ', as.character(qqq$condition_satz))

ggplot2::ggplot(data = qqq, aes(x = resid + 0.17, fill = condition_total)) +
  geom_density(alpha = .3, color = NA) +
  geom_rug(aes(x = SST.resid.SMB + 0.17, y = 0), position = position_jitter(height = 0)) +
  ggplot2::ggsave('band_diff_and_satz_empirical_density.pdf')

```

```{r empirical_density_x1, echo=TRUE}

qqq <- data.frame(condition = ifelse(orig$x1 > 10, ifelse(orig$x1 > 25, 2, 1), 0), resid = orig$SST.resid.SMB)
qqq$condition <- factor(qqq$condition, levels = c(0, 1, 2), labels = c('T_11 less than 10', 'T_11 in-between 10 and 25', 'T_11 greater than 25'))

ggplot2::ggplot(data = qqq, aes(x = resid + 0.17, fill = condition)) +
  geom_density(alpha = .4, color = NA) +
  geom_rug(aes(x = SST.resid.SMB + 0.17, y = 0), position = position_jitter(height = 0)) +
  ggsave('buoy_sst_empirical_density_environmental_context.pdf')
```

```{r empirical_density_x2, echo=TRUE}

cutoff_val <- 10
qqq <- data.frame(condition = ifelse(orig$x2 > cutoff_val, 1, 0), resid = orig$SST.resid.SMB)
qqq$condition <- factor(qqq$condition, levels = c(0, 1), labels = c('x2 less than or equal to 10', 'x2 greater than 10'))

ggplot2::ggplot(data = qqq, aes(x = resid, fill = condition)) +
  geom_density(alpha = .4, color = NA) +
  geom_rug(aes(x = SST.resid.SMB + 0.17, y = 0), position = position_jitter(height = 0)) +
  ggsave('x2_empirical_density.pdf')

```

```{r empirical_density_buoy_sst_environmental_context, echo=TRUE}

qqq <- data.frame(condition = ifelse(orig$buoy.sst > 10, ifelse(orig$buoy.sst > 25, 2, 1), 0), resid = orig$SST.resid.SMB)
qqq$condition <- factor(qqq$condition, levels = c(0, 1, 2), labels = c('buoy SST less than 10', 'buoy SST in-between 10 and 25', 'buoy SST greater than 25'))

ggplot2::ggplot(data = qqq, aes(x = resid, fill = condition)) +
  geom_density(alpha = .4, color = NA) +
  geom_rug(aes(x = SST.resid.SMB + 0.17, y = 0), position = position_jitter(height = 0)) +
  ggsave('buoy_sst_empirical_density_environmental_context.pdf')

```

```{r empirical_density_range_11, echo=TRUE}

cutoff_val <- 1
qqq <- data.frame(condition = ifelse(orig$range11 > cutoff_val, 1, 0), resid = orig$SST.resid.SMB)
qqq$condition <- factor(qqq$condition, levels = c(0, 1), labels = c('range11 less than or equal to 1', 'range11 greater than 1'))

ggplot2::ggplot(data = qqq, aes(x = resid, fill = condition)) +
  geom_density(alpha = .4, color = NA) +
  geom_rug(aes(x = SST.resid.SMB + 0.17, y = 0), position = position_jitter(height = 0)) +
  ggsave('range_11_empirical_density.pdf')

```

```{r empirical_density_product_buoy_SST_and_band_diff, echo=TRUE}

# Question: cutoff values for this plot?
prod_buoy_SST_band_diff <- orig$buoy.sst * orig$band.diff
qqq <- data.frame(condition = ifelse(prod_buoy_SST_band_diff > 10, ifelse(prod_buoy_SST_band_diff > 30, ifelse(prod_buoy_SST_band_diff > 75, 3, 2), 1), 0), resid = orig$SST.resid.SMB)
qqq$condition <- factor(qqq$condition, levels = c(0, 1, 2, 3), labels = c('Product less than 10',
                                                                       'Product in-between 10 and 30',
                                                                       'Product in-between 30 and 75',
                                                                       'Product greater than 75'))

ggplot2::ggplot(data = qqq, aes(x = resid, fill = condition)) +
  geom_density(alpha = .4, color = NA) +
  geom_rug(aes(x = SST.resid.SMB + 0.17, y = 0), position = position_jitter(height = 0)) +
  ggsave('product_buoy_SST_band_diff_empirical_density.pdf')

```



