---
title: "NLSST Exploration for AGU"
author: "Chirag Kumar and Guillermo Podesta"
output:
  html_notebook:
    toc: True
    theme: united
---

In this notebook, we explore the rationale behind the NLSST algorithm and where it fails. We investigate the
correlation between the temperature deficit and band difference. Finally, we aim to use decision trees to
estimate when a retrieval has a low or high residual magnitude based on the above exploration of when the
atmospheric correction is no longer an accurate proxy for the temperature deficit.


# Prep Workspace and Data
We use MODIS matchups from 2002 - 2016 that have been read in through a previous script. Here we simply load
the matchups, any required packages, and define necessary functions.

```{r prep_workspace, include=FALSE}
# Import necessary packages
require(ggplot2)
require(dplyr)
require(RColorBrewer)
require(circular)
require(ggmap)
require(raster)
require(sfsmisc)
require(rasterVis)
require(rgdal)
require(rpart)
require(rpart.plot)
require(caret)
require(evtree)
require(ctree)
require(partykit)
require(randomForest)
require(RWeka)
require(DMwR)
require(wesanderson)

ggplot <- function(...) {ggplot2::ggplot(...) + theme_bw()}

# Define secant function
secant.deg <- function(x) {1 / (cos(circular::rad(x)))}

# Source own functions
# Define direwctory where functions are for each operating system

if (Sys.info()["sysname"] == 'Windows') {
  fun.dir <- 'D:/matchups/r-projects/Matchup_R_Scripts/Functions/'
} else if (Sys.info()["sysname"] == 'Linux') {
  fun.dir <- '/home/ckk/Projects/Matchup_R_Scripts/Functions/'
}

fun.file <- paste0(fun.dir, 'common_functions.R')

source(file = fun.file,
  local = FALSE, echo = FALSE, verbose = FALSE)
rm(fun.dir, fun.file)


# Load data
# For 787
linux_dir <- '~/Projects/Matchup_R_Scripts/Results/objects/'
# For Laptop
#linux_dir <- '~/Projects/Matchup_R_Scripts/'
linux_file <- 'MODIS_Aqua_GSFC_ALL_Class_6.4.1_ao_2017_04_12_with_ancillary.Rdata'
AQUA_file <- paste0(linux_dir, linux_file)

if (!file.exists(AQUA_file)) {
    stop('Input file does not exist')
  } else {
    load(AQUA_file, verbose = TRUE)
    AQUA <- MODIS_Aqua_GSFC_ALL_Class_6.4.1_ao_2017_04_12
  }

rm(linux_dir, linux_file, AQUA_file)
rm(MODIS_Aqua_GSFC_ALL_Class_6.4.1_ao_2017_04_12)

set.seed(42)
```

The matchup file read in at a previous step is not filtered and contains too many variables. For the purpose of 
this analysis, we use only nighttime matchups with a quality of 0, 1, or 2 and only keep variables from the infrared
channels we use.

```{r prep_data, echo=TRUE}
# Turn POSIXct objects to characters bc dplyr doesn't support POSIXct
AQUA$sat.timedate <- as.character(AQUA$sat.timedate)
AQUA$buoy.timedate <- as.character(AQUA$buoy.timedate)

# Apply basic filtering
AQUA <- dplyr::tbl_df(AQUA) %>%
  dplyr::filter(solz >= 90) %>%
  dplyr::filter(qsst == 0 | qsst == 1 | qsst == 2)

# Now grab only variables that may be used to determine retrieval accuracy and make some new variables (i.e. x1, x2, x3)
# Create df of features
orig <- dplyr::tbl_df(AQUA) %>%
  dplyr::mutate(T_11 = cen.11000,
    T_12 = cen.12000,
    band.diff = T_11 - T_12,
    ref_SST = cen.ref.type.1.SST,
    x2 = band.diff * ref_SST,
    x3 = ((secant.deg(satz) - 1) * band.diff),
    lat = buoy.lat,
    lon = buoy.lon,
    sd11 = sd.11000,
    sd12 = sd.12000,
    range11 = max.11000 - min.11000,
    range12 = max.12000 - min.12000,
    diff.med.min11 = med.11000 - min.11000,
    diff.med.min12 = med.12000 - min.12000,
    qsst = qsst,
    buoy.sst = buoy.sst,
    cell5deg = cell5deg,
    buoy.timedate = buoy.timedate,
    SST.resid.SMB = cen.sst - (buoy.sst - 0.17)) %>% # SMB = sat minus buoy - also debias buoy.sst by turning buoy.sst into a skin measurement
  dplyr::select(T_11, T_12, band.diff, ref_SST, x2, x3, satz, lon, lat, sd11, sd12, range11, range12, diff.med.min11, diff.med.min12,
    qsst, buoy.sst, cell5deg, buoy.timedate, SST.resid.SMB)

```

# SST and Residuals

```{r residuals_vs_buoy_SST, echo = TRUE}

ggplot2::ggplot(data = orig, mapping = aes(x = buoy.sst, y = SST.resid.SMB,
  fill = cut(..count.., c(0, 10, 50, 100, 250, 500, 1000, 2500, 5000, Inf)))) +
  geom_bin2d(bins = 75) +
  scale_fill_hue('count') +
  geom_abline(slope = 0, intercept = 0, col = 'black')

```
We see that as the SST increases, so does the spread of the residuals. There are also more retrievals at higher
SSTs. We now explore the NLSST algorithm to find why there is an increase in residual spread at higher SSTs.

## Residual Distribution

```{r resid_distribution, echo=TRUE}
resid_cats <- cut(orig$SST.resid.SMB, c(-Inf, -0.4, 0.4, Inf))
table(resid_cats)
```

Because of the very few retrievals that have residuals > 0.4, we choose to discard these from our analysis.
Furthermore, these residuals are only causes by extremely abnormal conditions and are hence easier to
identify. For ease, we refer to retrievals less than -0.4 as bad and retrievals inbetween -0.4 and 0.4 as
good.

```{r resid_filtering, echo=TRUE}
orig <- dplyr::tbl_df(orig) %>%
  dplyr::filter(SST.resid.SMB <= 0.4)

orig <- as.data.frame(orig)
summary(orig$SST.resid.SMB)

resid_cats <- cut(orig$SST.resid.SMB, c(-Inf, -0.4, 0.4, Inf))
table(resid_cats)
```

After the series of filters that we have applied, we are left with `r nrow(orig)` matchups, ranging from
`r min(orig$buoy.timedate)` to `r max(orig$buoy.timedate)`

# NLSST Exploration

```{r band_diff_temp_deficit_qsst_all, echo=TRUE}

ccc = RColorBrewer::brewer.pal(n = 8, 'YlOrRd')
ccc <- ccc[3:8]

temp_deficit <- orig$buoy.sst - orig$T_11

# Take a sample of orig bc loess requires lots of memory - for testing purposes
index <- sample(1:nrow(orig), 50000, replace = FALSE)
zz1 <- zz1[index, ]

ggplot2::ggplot(mapping = aes(x = (orig$band.diff[index] * orig$ref_SST[index]), y = temp_deficit[index])) +
  geom_bin2d(bins = 20) +
  scale_fill_gradientn(colours = ccc) +
  geom_smooth(method = lm, se = FALSE, color = 'green', alpha = 0.1) +
  geom_smooth(method = loess, se = FALSE, span = 0.4) +
  labs(x = '(T_11 - T_12) * Reference SST', y = 'Temperature Deficit')
```

At lower SSTs, there is a strong linear correlation between the product and the temperature deficit,
indicating that at low SSTs, the NLSST atmospheric correction works very well.

We also see that at higher SSTs, there is a breakdown of the linear correlation between the temperature
deficit and product used in the NLSST algorithm (x_2). The temperature deficit increases more than
the band difference does, which causes increase in retrievals with negative residuals at higher
temperatures. However, the band difference alone is not enough for us to separate retrievals
with good and bad residual ranges.

```{r band_diff_temp_deficit_context_residual_qsst0_only, echo=TRUE}

resid_context <- cut(orig$SST.resid.SMB, c(-Inf, -.4, .4, Inf))
temp_deficit <- orig$buoy.sst - orig$T_11

uuu <- data.frame(resid_context = as.factor(resid_context), temp_deficit = temp_deficit, product = (orig$band.diff * orig$ref_SST))

# Take a sample of uuu bc loess requires lots of memory - for testing purposes
set.seed(42)
index <- sample(1:nrow(uuu), 50000, replace = FALSE)
uuu <- uuu[index, ]


ggplot2::ggplot(data = uuu, mapping = aes(x = product, y = temp_deficit, color = uuu$resid_context)) +
  geom_density2d() +
  labs(x = '(T_11 - T_12) * Reference SST', y = 'Temperature Deficit', color = 'Residual Range')
  #ggsave('band_diff_predictor_of_temp_deficit_different_in_situ_windows.pdf')

```



```{r median_by_location_raster, echo=TRUE}
# --- These numbers may help perform statistics for the matchups
# --- (e.g., number of matchups per cell, etc.).


resid_context <- cut(orig$SST.resid.SMB, c(-Inf, -.4, .4, Inf))
temp_deficit <- orig$buoy.sst - orig$T_11

uuu <- data.frame(resid_context = as.factor(resid_context), temp_deficit = temp_deficit, product = (orig$band.diff * orig$ref_SST),
                  SST.resid.SMB = orig$SST.resid.SMB)

# Take a sample of uuu bc loess requires lots of memory - for testing purposes
set.seed(42)
index <- sample(1:nrow(uuu), 50000, replace = FALSE)
uuu <- uuu[index, ]

# --- Create a raster object with 5-degree pixels

grid5deg <- raster::raster(ncol = 20, nrow = 40,
  xmn = min(uuu$product), xmx = max(uuu$product),
  ymn = min(uuu$temp_deficit), ymx = max(uuu$temp_deficit))
  #crs = crs.string)

# --- Write cell numbers as value for the grids.
# --- Cell 1 is the upper left corner and numbers go
# --- across the top line through its end, then
# --- start again in the second line, and so forth.

raster::values(grid5deg) <- 1:raster::ncell(grid5deg) # Cell numbers for 5 deg grid

raster::cellFromRowCol(object = grid5deg, rownr = 2, colnr = 20)

raster::xyFromCell(object =  grid5deg, cell = 800, spatial=FALSE)

pts <- uuu
sp::coordinates(pts) <- ~product + temp_deficit
  
cell5deg <- raster::rasterize(pts, grid5deg, pts$SST.resid.SMB, fun = median)

#qqq <- dplyr::tbl_df(uuu) %>%
#  dplyr::group_by(cell5deg) %>%
#  dplyr::summarise(N = n(), med = median(SST.resid.SMB), IQR = IQR(SST.resid.SMB)) %>%
#  dplyr::filter(N >= 100)

#qq1 <- rep(NA, (20*40))
#qq1[qqq$cell5deg] <- qqq$med

#raster::values(grid5deg) <- qq1

blues_palette <- rev(RColorBrewer::brewer.pal(9, 'Blues'))
more_blues <- colorRampPalette(blues_palette)(13)
reds_palette <- RColorBrewer::brewer.pal(6, 'Reds')

palette <- c(more_blues, '#f0f0f0', '#f0f0f0', reds_palette)

rasterVis::levelplot(cell5deg, margin = FALSE,
  at = c(-Inf, -.8, -.7, -.6, -.5, -.4, -.3, -.2, -.1, 0, 0.1, 0.2, 0.3, Inf),
  col.regions = palette)
```






